---
title: "SI_relative_v08"
author: "Asbjørn"
date: "3/9/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
library(readxl)
library(edgeR) #normalisation and DE analysis
library(RColorBrewer) 
library(gridExtra) #writing tables + arranging plots on pages
library(performance) #for summarising model performance
library(rcompanion) #for nagelkerke
library(pROC) #(in order to get confidence interval of the AUC)
library(glmnet) #lasso modelling
library(rstatix) #for t-tests etc in pipe, also multiple testing adjustments.
library(annotables) #simplified conversion of gene names.
library(scales)
#library(conflicted)
#conflict_prefer("filter", "dplyr")


theme_set(theme_bw()) #change default ggplot background
theme_update(text= element_text(size = 7, family = "sans"))
#import self-made functions
source("/Users/asbjorh/PhD/RNAseq_temporal/analysis/OFC/07_common_scripts/functions.R")
cols <- scales::viridis_pal()(3)[c(3,2)] #yellow and green
high_L2G <- read_tsv("/Users/asbjorh/PhD/RNAseq_temporal/genes_of_interest/OpenTargets/genes_l2g_0.5_locus.txt")
high_L2G_closest <- read_tsv("/Users/asbjorh/PhD/RNAseq_temporal/genes_of_interest/OpenTargets/loci_missing_nearest.txt")
high_L2G_closest <- high_L2G_closest %>% rename("symbol"= Nearest_gene) %>%
bind_rows( select(high_L2G, symbol, Locus)) %>%  distinct()
subset <- high_L2G_closest$symbol
subset <- unique(subset)
subsetname <- "high_L2G_closest"
#genes_res_diff <- read_tsv("/Users/asbjorh/PhD/RNAseq_temporal/analysis/OFC/core_genes_consistent_res_diff.txt", col_names = FALSE)
#subset <- genes_res_diff$X1
#subsetname <- "genes_res_diff_22"

#the liability scale takes into account the discrepancy bewtween number of cases in the test set and in the population
#k <- 0.02 #according to Kevin O'Connell, this prevalence is what was used in the BD GWAS PGC3
#p <- 71/(167+71) #proportion cases

result_h2l <- function(k, r2n, p) {
x <- qnorm(1 - k)
z <- dnorm(x)
i <- z / k
cc <- k * (1 - k) * k * (1 - k) / (z^2 * p * (1 - p))
theta <- i * ((p - k)/(1 - k)) * (i * ((p - k) / ( 1 - k)) - x)
e <- 1 - p^(2 * p) * (1 - p)^(2 * (1 - p))
h2l <- cc * e * r2n / (1 + cc * e * theta * r2n)
}


```


```{r edgeR normalisation}
dir <- "/Users/asbjorh/PhD/RNAseq_DE/"

Set <- "CMC-HBCC"
counts <- count_CM
gene.length <- read_tsv( paste0(dir, "gene.length.tsv")) # these are lengths from ensembl, either canonical transcript, or median length if no canonical is defined.
samples_Set <- samples[samples$Set==Set, 2]
info_Set <- info_all %>% filter(SampleID %in% samples_Set$SampleID)
info_Set <- info_Set[order(samples_Set$SampleID),]
counts_Set <- counts[, colnames(counts) %in% samples_Set$SampleID]
counts_Set <- counts_Set[,order(samples_Set$SampleID)]
counts_Set  <- counts_Set[(rownames(counts_Set ) %in% gene.length$gene_id),] #remove the counts for genes we don't have lengths for
#order gene lengths
counts_Set  <- counts_Set[,order(samples_Set$SampleID)]
gene.length <- gene.length %>% filter(gene_id %in% rownames(counts_Set))
gene.length <- gene.length[match(rownames(counts_Set ), gene.length$gene_id), ] #order the same as counts
setequal(rownames(counts_Set), gene.length$gene_id) #double check that genes and lengths match
#normalise with edgeR, and have edgeR do the expression-level filtering and log2-transformation
#edgeR differential expression
group <- info_Set$PrimaryDiagnosis
y <- DGEList(counts=counts_Set,
#group=group,
samples = info_Set,
genes= gene.length)
#library(edgeR)
y <- DGEList(counts=counts_Set,
#group=group,
samples = info_Set,
genes= gene.length)
keep <- filterByExpr(y)
y <- y[keep,]
y <- calcNormFactors(y)
rpkm_Set <- edgeR::rpkm(y,
normalized.lib.sizes = TRUE,
log = TRUE,
#log = FALSE #,
prior.count = 0.01
)
rpkm_Set <- as.data.frame(rpkm_Set)
#replace ensembl IDs with gene symbols
rpkm_Set <- cbind(symbol = grch38$symbol[ match(rownames(rpkm_Set), grch38$ensgene) ],
rpkm_Set)
rpkm_Set$symbol[duplicated(rpkm_Set$symbol)] <- NA #mark the duplicated gene names as NA
rpkm_Set <- tibble::as_tibble(rpkm_Set, rownames = "gene_id") %>%
mutate(symbol = coalesce(symbol, gene_id)) #insert ensembl id if NA
rpkm_Set <- as.data.frame(rpkm_Set[,-1]) #remove the ensemble ids
rownames(rpkm_Set) <- rpkm_Set$symbol #make the gene names rownames
rpkm_Set <- rpkm_Set[,-1] #remove the gene name column
rpkm_log_CMC_HBCC <- rpkm_Set


Set <- "CMC-Pitt"
counts <- count_CM
gene.length <- read_tsv( paste0(dir, "gene.length.tsv")) # these are lengths from ensembl, either canonical transcript, or median length if no canonical is defined.
samples_Set <- samples[samples$Set==Set, 2]
info_Set <- info_all %>% filter(SampleID %in% samples_Set$SampleID)
info_Set <- info_Set[order(samples_Set$SampleID),]
counts_Set <- counts[, colnames(counts) %in% samples_Set$SampleID]
counts_Set <- counts_Set[,order(samples_Set$SampleID)]
counts_Set  <- counts_Set[(rownames(counts_Set ) %in% gene.length$gene_id),] #remove the counts for genes we don't have lengths for
#order gene lengths
counts_Set  <- counts_Set[,order(samples_Set$SampleID)]
gene.length <- gene.length %>% filter(gene_id %in% rownames(counts_Set))
gene.length <- gene.length[match(rownames(counts_Set ), gene.length$gene_id), ] #order the same as counts
setequal(rownames(counts_Set), gene.length$gene_id) #double check that genes and lengths match
#normalise with edgeR, and have edgeR do the expression-level filtering and log2-transformation
#edgeR differential expression
group <- info_Set$PrimaryDiagnosis
y <- DGEList(counts=counts_Set,
#group=group,
samples = info_Set,
genes= gene.length)
keep <- filterByExpr(y)
y <- y[keep,]
y <- calcNormFactors(y)
rpkm_Set <- edgeR::rpkm(y,
normalized.lib.sizes = TRUE,
log = TRUE,
#log = FALSE #,
prior.count = 0.01
)
rpkm_Set <- as.data.frame(rpkm_Set)
#replace ensembl IDs with gene symbols
rpkm_Set <- cbind(symbol = grch38$symbol[ match(rownames(rpkm_Set), grch38$ensgene) ],
rpkm_Set)
rpkm_Set$symbol[duplicated(rpkm_Set$symbol)] <- NA #mark the duplicated gene names as NA
rpkm_Set <- tibble::as_tibble(rpkm_Set, rownames = "gene_id") %>%
mutate(symbol = coalesce(symbol, gene_id)) #insert ensembl id if NA
rpkm_Set <- as.data.frame(rpkm_Set[,-1]) #remove the ensemble ids
rownames(rpkm_Set) <- rpkm_Set$symbol #make the gene names rownames
rpkm_Set <- rpkm_Set[,-1] #remove the gene name column
# rpkm_log_CMC_HBCC <- rpkm_Set
rpkm_log_CMC_Pitt <- rpkm_Set


Set <- "GVEX-SMRI"
counts <- count_GVEX
gene.length <- read_tsv( paste0(dir, "gene.length.tsv")) # these are lengths from ensembl, either canonical transcript, or median length if no canonical is defined.
samples_Set <- samples[samples$Set==Set, 2]
info_Set <- info_all %>% filter(SampleID %in% samples_Set$SampleID)
info_Set <- info_Set[order(samples_Set$SampleID),]
counts_Set <- counts[, colnames(counts) %in% samples_Set$SampleID]
counts_Set <- counts_Set[,order(samples_Set$SampleID)]
counts_Set  <- counts_Set[(rownames(counts_Set ) %in% gene.length$gene_id),] #remove the counts for genes we don't have lengths for
#order gene lengths
counts_Set  <- counts_Set[,order(samples_Set$SampleID)]
gene.length <- gene.length %>% filter(gene_id %in% rownames(counts_Set))
gene.length <- gene.length[match(rownames(counts_Set ), gene.length$gene_id), ] #order the same as counts
setequal(rownames(counts_Set), gene.length$gene_id) #double check that genes and lengths match
#normalise with edgeR, and have edgeR do the expression-level filtering and log2-transformation
#edgeR differential expression
group <- info_Set$PrimaryDiagnosis
y <- DGEList(counts=counts_Set,
#group=group,
samples = info_Set,
genes= gene.length)
keep <- filterByExpr(y)
y <- y[keep,]
y <- calcNormFactors(y)
rpkm_Set <- edgeR::rpkm(y,
normalized.lib.sizes = TRUE,
log = TRUE,
#log = FALSE #,
prior.count = 0.01
)
rpkm_Set <- as.data.frame(rpkm_Set)
#replace ensembl IDs with gene symbols
rpkm_Set <- cbind(symbol = grch38$symbol[ match(rownames(rpkm_Set), grch38$ensgene) ],
rpkm_Set)
rpkm_Set$symbol[duplicated(rpkm_Set$symbol)] <- NA #mark the duplicated gene names as NA
rpkm_Set <- tibble::as_tibble(rpkm_Set, rownames = "gene_id") %>%
mutate(symbol = coalesce(symbol, gene_id)) #insert ensembl id if NA
rpkm_Set <- as.data.frame(rpkm_Set[,-1]) #remove the ensemble ids
rownames(rpkm_Set) <- rpkm_Set$symbol #make the gene names rownames
rpkm_Set <- rpkm_Set[,-1] #remove the gene name column
rpkm_log_GVEX_SMRI <- rpkm_Set
rpkm_log_GVEX_SMRI[1:6,1:6]


Set <- "BipSeq-sACC"
counts <- count_BipSeq
gene.length <- read_tsv( paste0(dir, "gene.length.tsv")) # these are lengths from ensembl, either canonical transcript, or median length if no canonical is defined.
samples_Set <- samples[samples$Set==Set, 2]
info_Set <- info_all %>% filter(SampleID %in% samples_Set$SampleID)
info_Set <- info_Set[order(samples_Set$SampleID),]
counts_Set <- counts[, colnames(counts) %in% samples_Set$SampleID]
counts_Set <- counts_Set[,order(samples_Set$SampleID)]
counts_Set  <- counts_Set[(rownames(counts_Set ) %in% gene.length$gene_id),] #remove the counts for genes we don't have lengths for
#order gene lengths
counts_Set  <- counts_Set[,order(samples_Set$SampleID)]
gene.length <- gene.length %>% filter(gene_id %in% rownames(counts_Set))
gene.length <- gene.length[match(rownames(counts_Set ), gene.length$gene_id), ] #order the same as counts
setequal(rownames(counts_Set), gene.length$gene_id) #double check that genes and lengths match
#normalise with edgeR, and have edgeR do the expression-level filtering and log2-transformation
#edgeR differential expression
group <- info_Set$PrimaryDiagnosis
y <- DGEList(counts=counts_Set,
#group=group,
samples = info_Set,
genes= gene.length)
keep <- filterByExpr(y)
y <- y[keep,]
y <- calcNormFactors(y)
rpkm_Set <- edgeR::rpkm(y,
normalized.lib.sizes = TRUE,
log = TRUE,
#log = FALSE #,
prior.count = 0.01
)
rpkm_Set <- as.data.frame(rpkm_Set)
#replace ensembl IDs with gene symbols
rpkm_Set <- cbind(symbol = grch38$symbol[ match(rownames(rpkm_Set), grch38$ensgene) ],
rpkm_Set)
rpkm_Set$symbol[duplicated(rpkm_Set$symbol)] <- NA #mark the duplicated gene names as NA
rpkm_Set <- tibble::as_tibble(rpkm_Set, rownames = "gene_id") %>%
mutate(symbol = coalesce(symbol, gene_id)) #insert ensembl id if NA
rpkm_Set <- as.data.frame(rpkm_Set[,-1]) #remove the ensemble ids
rownames(rpkm_Set) <- rpkm_Set$symbol #make the gene names rownames
rpkm_Set <- rpkm_Set[,-1] #remove the gene name column
rpkm_log_BipSeq_sACC <- rpkm_Set

```


```{r glmnet lasso model functions}
#######function to build models #########
lasso_model <- function(modelset, GOI, info){
  modelset_t <- as.data.frame(t(modelset)) 
 
   #insert covariates age, sex, RIN, PMI , ethnicity where applicable
  
modelset_t <- cbind(SEX = info$ReportedGender[match(rownames(modelset_t), info$SampleID)], modelset_t)

modelset_t <- cbind(AGE = info$Age[match(rownames(modelset_t), info$SampleID)], modelset_t)
modelset_t <- cbind(RIN = info$RIN[match(rownames(modelset_t), info$SampleID)], modelset_t)
modelset_t <- cbind(PMI = info$PMI[match(rownames(modelset_t), info$SampleID)], modelset_t)
modelset_t <- cbind(Ethnicity = info$Ethnicity[match(rownames(modelset_t), info$SampleID)], modelset_t) #remove for BipSeq

#dummy Sex:
modelset_t$Male <- ifelse(modelset_t$SEX == "Male", 1, 0)

#dummy Ethnicity
modelset_t$ethn_cauc <-  ifelse(modelset_t$Ethnicity == "Caucasian", 1, 0)

modelset_t <- modelset_t %>% select(
  -Ethnicity, 
  -SEX)
modelset_t <- modelset_t %>% relocate(Male, 
                                      ethn_cauc, 
                                      .after= AGE)

modelset_t <- as.matrix(modelset_t)

# nested for-loop for each gene against each other gene
  lasso_models <- list() # Create empty lists
  lasso_beta <- list()
    for( gene in GOI){
#first-pass to find the optimal lambda
      lasso.cv.tr <- cv.glmnet(x=modelset_t[,!colnames(modelset_t) %in% gene], y = modelset_t[,colnames(modelset_t) %in% gene], 
                               alpha=1,
                               type.measure="deviance",
                               penalty.factor = p.fac, 
                         family="gaussian", standardize=TRUE, nfolds=10) #remove the gene being modelled from the data in the model
  #    lasso.lam.tr <- lasso.cv.tr$lambda.min #optimal lambda (minimal)
      lasso.lam.tr <- lasso.cv.tr$lambda.1se #optimal lambda 1se
      lasso.mod.tr <- lasso.cv.tr$glmnet.fit #extract the fitted models
      # Extract the final model with the optimal lambda. Lasso:
      lasso.final.tr <- glmnet(x=modelset_t[,!colnames(modelset_t) %in% gene], y =  modelset_t[,colnames(modelset_t) %in% gene], family="gaussian",  alpha=1, 
                              penalty.factor = p.fac,   #this is where some variables are fixed, and not allowed to be penalised. p.fac is set to 0 (default 1)
                               lambda=lasso.lam.tr
                               )
      lasso_models[[gene]] <- lasso.final.tr
      }
lasso_models
}


####function to get residuals in the testset#####
lasso_test <- function(testset, models, GOI, info){
  testset_t <- as.data.frame(t(testset)) 
  #insert covariates age, sex, RIN, PMI  
testset_t <- cbind(SEX = info$ReportedGender[match(rownames(testset_t), info$SampleID)], testset_t)

testset_t <- cbind(AGE = info$Age[match(rownames(testset_t), info$SampleID)], testset_t)
testset_t <- cbind(RIN = info$RIN[match(rownames(testset_t), info$SampleID)], testset_t)
testset_t <- cbind(PMI = info$PMI[match(rownames(testset_t), info$SampleID)], testset_t)
testset_t <- cbind(Ethnicity = info$Ethnicity[match(rownames(testset_t), info$SampleID)], testset_t) #remove for BipSeq

#dummy Sex:
testset_t$Male <- ifelse(testset_t$SEX == "Male", 1, 0)

#dummy Ethnicity
testset_t$ethn_cauc <-  ifelse(testset_t$Ethnicity == "Caucasian", 1, 0)

testset_t <- testset_t %>% select(
  -Ethnicity, 
  -SEX)
testset_t <- testset_t %>% relocate(Male, 
                                    ethn_cauc, 
                                    .after= AGE)


#for each gene in the testset, predict it based on the model and store each individual's residuals
lasso_pred = array(dim=c(length(GOI), length(rownames(testset_t))))
R2= c() #create vector for the R2 
rownames(lasso_pred) <- GOI
colnames(lasso_pred) <- rownames(testset_t)
  for( gene in GOI){
    test <- testset_t[,!colnames(testset_t) %in% gene]
    test <- as.matrix(test)
    pred =  predict(models[[gene]], s=models[[gene]]$lambda, type="response", newx=test)
    lasso_pred[gene,] <- pred 
    R2[gene] <- models[[gene]]$dev.ratio
  }

testset_obs <- testset_t[,colnames(testset_t) %in% GOI] 
testset_obs <- t(testset_obs)
testset_obs <- testset_obs[match(rownames(lasso_pred),rownames(testset_obs)), 
                           match(colnames(lasso_pred),colnames(testset_obs))] #reorder the observed to match the predicted. Keep only matching, i.e. remove the covariate values
lasso_residuals <- testset_obs - lasso_pred #create a matrix with all residuals
lasso_residuals_std <-  lasso_residuals/mean(testset_obs)  #adjust the residuals to the mean of observed values
lasso_residuals_std
}




```

```{r gene subset}
# Get one gene per GWAS locus
subset <- high_L2G_closest$symbol


subset <- unique(subset) # because 2 loci share a gene.
length(subset)
```


Prepare the expression and covariate data
```{r}
info_all <- info_all %>%  filter(!SampleID %in% remove_PCA)

dataset <- "CMC-HBCC"
dataset <- "CMC-Pitt"
dataset <- "BipSeq-sACC"
dataset <- "GVEX-SMRI"

info <- info_all %>%  filter(Set== dataset) 

  
rpkm_log_CMC_HBCC <- rpkm_log_CMC_HBCC[,!colnames(rpkm_log_CMC_HBCC) %in% remove_PCA]
rpkm_log <- rpkm_log_CMC_HBCC 

 rpkm_log <- rpkm_log_CMC_Pitt
 rpkm_log <- rpkm_log_BipSeq_sACC
 rpkm_log <- rpkm_log_GVEX_SMRI

subset <- subset[subset %in% rownames(rpkm_log)]

HC <- info %>% filter(PrimaryDiagnosis=="HC") %>% select(SampleID)
BD <-  info %>% filter(PrimaryDiagnosis=="BD") %>% select(SampleID)

setequal(info$SampleID, colnames(rpkm_log))
```



Model and calculate residuals

```{r run iterations of lasso test}
debug = TRUE

it <- 100 #number of iterations  
#it <- 300 #for BrainGVEX_SMRI
#it <- 1000 #for BrainGVEX_SMRI

model_fraction <- length(HC$SampleID) - length(BD$SampleID) #for CMC_HBCC
#model_fraction <- length(HC$SampleID) *0.9 #for BrainGVEX_SMRI
#model_fraction <- length(HC$SampleID) *3/4 #for CMC_Pitt
#model_fraction <- length(HC$SampleID) *2/3 #for BipSeq sACC

#save the results from iterations to list
res_std_C_iterate <- list()
res_std_BD_iterate<- list()
R2_iterate <- list()  
models_iterate <- list()  
#beta_iterate <- list()
#function to save named beta
#?


#Which variables (covarites) should be fixed?
#for CMC, GVEX where ethnicity differs 
   p.fac = rep(1, length(subset) -1 + 5) 
    p.fac[1:5] = 0


#for BipSeq where ethnicity does not differ
#    p.fac = rep(1, length(subset) -1 + 4) 
#    p.fac[1:4] = 0
    
  #randomly picked sets of controls for the modeling
for(i in 1:it){
  set.seed(i) 
  HC_model <- slice_sample(HC, n= model_fraction) 
  modelset <- rpkm_log[rownames(rpkm_log) %in% subset,colnames(rpkm_log) %in% HC_model$SampleID]
  testset1 <- rpkm_log[rownames(rpkm_log) %in% subset,colnames(rpkm_log) %in% HC$SampleID & !colnames(rpkm_log) %in% HC_model$SampleID]
  testset2 <- rpkm_log[rownames(rpkm_log) %in% subset,colnames(rpkm_log) %in% BD$SampleID]
  lasso_models <- lasso_model(modelset = modelset, GOI = subset, info =info) #these functions are defined in the chunk "multiple linear model all GOI functions"
  R2 <- sapply(lasso_models, function(x) as.vector(x[["dev.ratio"]]))
  test_C <- lasso_test(testset = testset1, models  = lasso_models , GOI = subset, info =info )
  test_BD <- lasso_test(testset = testset2, models = lasso_models , GOI = subset , info =info)
  models_iterate[[i]] <- lasso_models 
  res_std_C_iterate[[i]] = test_C 
  res_std_BD_iterate[[i]] =  test_BD
  R2_iterate[[i]] = R2 #save the R2 from all models from the iteration
  
    #needs sanity check. Is the length of rownames(test_C) == rownames(test_BD) == R2_iterate[[i]]. Stop if not "all genes are not modelled. check gene subset for duplicates".
  # Modulus operation
  if(i %% 10==0) {
      cat(paste0("iteration: ", i, "\n"))
  }
  
  if(debug==TRUE & i == 10) {
      cat("HC res \n")
      print(matrix(test_C[1:4,1:4], nrow= 4, ncol=4, byrow = FALSE)) 
      print(colnames(test_C)[1:4])
      print(rownames(test_C)[1:4])
      cat("\nHC res has dimensions: \n")
      print(dim(test_C))
      cat("\nBD res \n")
      print(matrix(test_BD[1:4,1:4], nrow= 4, ncol=4, byrow = FALSE))
      print(colnames(test_BD)[1:4])
      print(rownames(test_BD)[1:4])
      cat("\nBD res has dimensions: \n")
      print(dim(test_BD))
      cat("\nR2 \n")
      cat(paste0(R2[1:4]), "\n")
      print(names(R2)[1:4])
      cat("number of R2:")
      print(length(R2))
  }
  
   if(debug==TRUE & i == 9) {
     cat("glimpse of the model dataset \n")
     print(modelset[1:8,1:8])
   }
  
}

```




```{r calculate WSR from residuals}

#weight residuals with R2
res_w_C_iterate <- lapply(1:length(res_std_C_iterate), function(x) res_std_C_iterate[[x]]*R2_iterate[[x]] ) #weight residuals according to R2 from the models: multiplication by the vector (R2) is by default done column-wise, since genes are in rows.
res_w_BD_iterate <- lapply(1:length(res_std_BD_iterate), function(x) res_std_BD_iterate[[x]]*R2_iterate[[x]] )


#melt function within list
meltfun <- function(x,it) {
   z <- reshape2::melt(x,id.vars=rownames(x))
  colnames(z) <- c('gene_id', 'SampleID', 'res')
  z$it = it
   return(z)
}


res_w_C_all_long <- mapply(meltfun,res_w_C_iterate, it= seq(1:length(res_w_C_iterate)),SIMPLIFY=FALSE) #melt each iteration to long
res_w_C_all<-  do.call("rbind", res_w_C_all_long) #combine all iterations into one dataframe

res_w_BD_all_long <- mapply(meltfun,res_w_BD_iterate, it= seq(1:length(res_w_BD_iterate)),SIMPLIFY=FALSE) #melt each iteration to long
res_w_BD_all<-  do.call("rbind", res_w_BD_all_long) #combine all iterations into one dataframe

#combine C and BD
res_w_all <- rbind(res_w_C_all, res_w_BD_all )

#calculate mean weighted residual across iterations
mean_res_w <- res_w_all %>%  
  group_by(gene_id, SampleID) %>%  
  summarise(mean_res_w= mean(res,  na.rm=TRUE)) %>% 
  ungroup()


```

```{r save R2}
#change directory
dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/", "fix_lambda/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_model/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "free_coeff_lasso/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "original_gene_selection/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "keep_TRPT1_LMAN2L/", dataset, "/WSR/")



models_info <- as.data.frame(do.call(cbind, R2_iterate)) #create a dataframe, each iteration in columns. 
models_info <- as_tibble(models_info, rownames = "gene_id") %>% rename_with(stringr::str_replace, #rename columns with iterations
              pattern = "V", replacement = "") %>% 
  pivot_longer(values_to = "R2", names_to = "iteration", cols = -1) %>% 
  group_by(gene_id) %>% mutate(gene_median= median(R2))

#write_tsv(models_info, paste0(dir, "models_info.txt"))

R2_models <- models_info %>%   
  ggplot(aes(x= reorder(gene_id, desc(gene_median)), #reorder according to gene median R2
             y=R2)) + 
    geom_boxplot(outlier.shape = NA) + 
    theme(text = element_text( size = 12 )) +
    theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust=1, size= 10)) +
  xlab("Gene modelled") +
  ylab("R2 in 100 iterations")

  ggsave(plot = R2_models, path = dir, filename = "R2_models.pdf", height = 12, width= 30, units="cm")
  
  
  
  
   #You can calculate R2 for the test data yourself, by extracting the deviance and null deviance and calculating dev.ratio=1-dev/nulldev.
```

Calculations based on WSR 


```{r calculate SI score}
#change directory
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_model/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda_model/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "free_coeff_lasso/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "original_gene_selection/", dataset, "/WSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "keep_TRPT1_LMAN2L/", dataset, "/WSR/")

#calculate SI
SI <- mean_res_w %>% 
  group_by(SampleID) %>% 
  summarise(SI = sum(abs(mean_res_w))/n_distinct(gene_id)) %>% #use the mean weighted scaled residual value 
  ungroup()
  
SI <- SI %>% left_join(select(info, SampleID, PrimaryDiagnosis), by= "SampleID")


SI <- rename(SI, "diag" = PrimaryDiagnosis)

```


```{r compare C and BD }
SI_C_BD <- SI

 options(digits = 5)
SI_HC <- filter(SI_C_BD, diag=="HC") 
SI_BD <- filter(SI_C_BD,  diag=="BD")
ks.test( SI_BD$SI,  SI_HC$SI)
kstest <- ks.test( SI_BD$SI,  SI_HC$SI)


```


```{r logistic regression}

SI$diag <- str_replace(SI$diag, pattern = "HC", replacement = "0")
SI$diag <- str_replace(SI$diag, pattern = "BD", replacement = "1")
SI$diag <- as.factor(SI$diag)

#SI <-  cbind(Ethnicity_code = info$Ethnicity_code[match(SI$SampleID, info$SampleID)], SI)
SI <-  cbind(Sex = info$ReportedGender[match(SI$SampleID, info$SampleID)], SI)
SI <-  cbind(Age = info$Age[match(SI$SampleID, info$SampleID)], SI)
#SI <-  cbind(Study = info$Study[match(SI$SampleID, info$SampleID)], SI) #For BrainGVEX


#dummy Sex:
SI$Male <- ifelse(SI$Sex == "Male", 1, 0)

SI <-  cbind(Ethnicity = info$Ethnicity[match(SI$SampleID, info$SampleID)], SI)
#dummy Ethnicity
SI$ethn_cauc <-  ifelse(SI$Ethnicity == "Caucasian", 1, 0)


SI_glm <- glm(diag ~ SI, data= as.data.frame(SI), family = "binomial")
#SI_glm_cov <- glm(diag ~ SI + Ethnicity_code + Sex + Age,
       #         +Study, #For BrainGVEX
 #                 data= as.data.frame(SI), family = "binomial") #check to see which variables effect outcome

SI_glm_cov <- glm(diag ~ SI  + Age + ethn_cauc + Male,
       #         +Study, #For BrainGVEX
                  data= as.data.frame(SI), family = "binomial") #check to see which variables effect outcome


summary <- summary(SI_glm)
summary_cov <- summary(SI_glm_cov)
perf <- model_performance(SI_glm)

#show the distribution of SI values as points
SI_boxplot <- SI %>% 
  rename(Diagnosis = diag) %>% 
 ggplot( aes(y=SI, x=Diagnosis, colour= Diagnosis))+ 
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(size=0.2) +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC", "BD")) +
  labs(colour="Diagnosis") + 
  xlab("Diagnosis")+ 
  theme(legend.key.size = unit(0.5,"line"),
        legend.position = ""
        ) + 
  scale_x_discrete(breaks=c(0,1),
                    labels=c("HC", "BD")) +
  ylab("SI score")

SI_ttest <- t.test(SI ~ diag, data = SI)
SI_wilcox <- wilcox.test(SI ~ diag, data = SI)

#get pseudo R2 from the model
pseudoR2 <- nagelkerke(SI_glm)


 r2l <- result_h2l(k=0.02, p=length(BD$SampleID)/(length(HC$SampleID) + length(BD$SampleID)), r2n = pseudoR2$Pseudo.R.squared.for.model.vs.null[3] #take out the Nagelkerke
                   ) 
```

```{r save analysis }
write_tsv(SI, file = paste0(dir,"SI_C_BD_wsr.txt"))
write.csv(summary$coefficients, file=paste0(dir, "glm_summary.txt"))
write.csv(summary_cov$coefficients, file=paste0(dir, "glm_covariates_summary.txt"))
write.csv(perf, file=paste0(dir, "glm_perf.txt"))
write.csv(pseudoR2$Pseudo.R.squared.for.model.vs.null, file=paste0(dir, "glm_pseudoR2.txt"))
write.csv( unlist(SI_ttest), file=paste0(dir, "SI_ttest.txt"))
write.csv( unlist(SI_wilcox), file=paste0(dir, "SI_wilcox.txt"))
write.csv(r2l, file=paste0(dir, "nagelkerke_liab.txt"))
write.csv(as.data.frame(unlist(kstest)), file = paste0(dir,"SI_kstest.txt"))
ggsave(plot= SI_boxplot, filename = paste0(dir, "SI_boxplot.pdf"), width = 43, height = 43, units = "mm")
```


Inspect the regression by ROC curves 
```{r AUC of the regression}
#ROC (Receiver operating characteristic) curve
pROC_obj <- roc(SI$diag,SI$SI,
                direction= "<", #when cases have lower SI, we need to force the direction
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)
sens.ci <- ci.se(pROC_obj)


sens_ci_fix <- as_tibble(sens.ci, rownames = "spec")
sens_ci_fix$spec <- as.numeric(sens_ci_fix$spec)

pROC <- 
  ggroc(pROC_obj, size=0.5) + 
  geom_abline(slope=1, intercept = 1, linetype = "dashed", alpha=0.7, color = "black") +
  geom_ribbon(data = sens_ci_fix, 
              aes(x = spec, ymin = `2.5%`, ymax = `97.5%`), 
              fill = "#053061", alpha= 0.2) +
  xlab("Specificity") + ylab("Sensitivity")

write.csv(pROC_obj$auc, file=paste0(dir, "auc_SI_WSR.txt"))
ggsave(pROC, filename = paste0(dir, "ROC_WSR.pdf"), width = 43, height = 43, units = "mm")  
```

Calculations based on sWSR (i.e. difference from HC mean)
```{r calculate sWSR}

#change directory

dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/", "fix_lambda/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda_model/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_model/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "free_coeff_lasso/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "original_gene_selection/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "keep_TRPT1_LMAN2L/", dataset, "/sWSR/")

#get diagnosis back in
mean_res_w_info <- mean_res_w %>% left_join(select(info, SampleID, PrimaryDiagnosis), by= "SampleID")
mean_res_w_info <- mean_res_w_info %>%  rename("wsr"= mean_res_w)

#calculate mean and sd for the HC
wsr_mean_HC <- mean_res_w_info %>% 
  filter(PrimaryDiagnosis =="HC") %>% 
  group_by(gene_id) %>% 
  mutate(mean_HC = mean(wsr),
         sd_HC = sd(wsr)) %>% 
  ungroup() %>% 
  select(gene_id, mean_HC, sd_HC) %>% 
  distinct()

#calculate sWSR-value. This is at first how many sd(HC) are the individual value from the mean HC?

sWSR_Set <- mean_res_w_info %>% 
  left_join(wsr_mean_HC, by="gene_id") %>% 
  mutate(sWSR = (wsr- mean_HC)/ sd_HC)

```


```{r SI score sWSR}
#calculate SI
SI_sWSR <- sWSR_Set %>% 
 drop_na() %>%  #when all models have R2=0 for a gene, then res will be 0 and SI NA
  group_by(SampleID) %>% 
  mutate(SI = sum(abs(sWSR))/n_distinct(gene_id)) %>% #use the mean weighted scaled residual value 
  ungroup() %>% 
  select(SampleID, PrimaryDiagnosis, SI) %>%
  distinct()

SI_sWSR<- rename(SI_sWSR, "diag" = PrimaryDiagnosis)

SI_sWSR_boxplot <- 
  SI_sWSR %>% 
  rename(Diagnosis = diag) %>% 
 ggplot( aes(y=SI, x=Diagnosis, colour= Diagnosis))+ 
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(size=0.2) +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC", "BD")) +
  labs(colour="Diagnosis") + 
  xlab("Diagnosis")+ 
  theme(legend.key.size = unit(0.5,"line"),
        legend.position = ""
        ) + 
  scale_x_discrete(labels=c("HC", "BD")) +
     scale_y_continuous(
       labels = label_number(accuracy = 0.1)
       ) +
  ylab("SI score")

SI_sWSR_dens <- 
  SI_sWSR %>% 
  ggplot(aes(x=SI, colour=diag)) + 
  geom_density(size=1) +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC", "BD")) +
  theme(legend.position = "none") +
  labs(x= "SI score", y= "Density")

ggsave(plot= SI_sWSR_boxplot, filename = paste0(dir, "SI_sWSR_boxplot.pdf"), width = 43, height = 43, units = "mm")
ggsave(plot = SI_sWSR_dens, filename = paste0(dir, "SI_density.pdf"), width = 43, height = 43, units = "mm")

```

```{r calculate ttest nagelkerke and auc}

SI_sWSR_ttest <- t.test(SI ~ diag, data = SI_sWSR)
SI_sWSR_wilcox <- wilcox.test(SI ~ diag, data = SI_sWSR)

SI_sWSR$diag <- str_replace(SI_sWSR$diag, pattern = "HC", replacement = "0")
SI_sWSR$diag <- str_replace(SI_sWSR$diag, pattern = "BD", replacement = "1")
SI_sWSR$diag <- as.factor(SI_sWSR$diag)

#SI_sWSR<-  cbind(Ethnicity_code = info$Ethnicity_code[match(SI_sWSR$SampleID, info$SampleID)], SI_sWSR)
SI_sWSR<-  cbind(Sex = info$ReportedGender[match(SI_sWSR$SampleID, info$SampleID)], SI_sWSR)
SI_sWSR<-  cbind(Age = info$Age[match(SI_sWSR$SampleID, info$SampleID)], SI_sWSR)
#SI_sWSR<-  cbind(Study = info$Study[match(SI_sWSR$SampleID, info$SampleID)], SI_sWSR) #For CMC-Pitt

#dummy Sex:
SI_sWSR$Male <- ifelse(SI_sWSR$Sex == "Male", 1, 0)

SI_sWSR <-  cbind(Ethnicity = info$Ethnicity[match(SI_sWSR$SampleID, info$SampleID)], SI_sWSR)
#dummy Ethnicity
SI_sWSR$ethn_cauc <-  ifelse(SI_sWSR$Ethnicity == "Caucasian", 1, 0)


SI_sWSR_glm <- glm(diag ~ SI, data= as.data.frame(SI_sWSR), family = "binomial")
#SI_sWSR_glm_cov <- glm(diag ~ SI + Sex + Age+ Ethnicity_code,
       #         +Study  , #For BrainGVEX
 #                 data= as.data.frame(SI_sWSR), family = "binomial") #check to see which variables effect outcome

SI_sWSR_glm_cov <- glm(diag ~ SI  + Age + ethn_cauc + Male,
       #         +Study, #For BrainGVEX
                  data= as.data.frame(SI_sWSR), family = "binomial") #check to see which variables effect outcome



summary_sWSR<- summary(SI_sWSR_glm)
summary_cov_sWSR<- summary(SI_sWSR_glm_cov)
perf_sWSR<- model_performance(SI_sWSR_glm)


#get pseudo R2 from the model
pseudoR2_sWSR<- nagelkerke(SI_sWSR_glm)
r2l_sWSR <- result_h2l(k=0.02, p=length(BD$SampleID)/(length(HC$SampleID) + length(BD$SampleID)), r2n = pseudoR2_sWSR$Pseudo.R.squared.for.model.vs.null[3] #take out the Nagelkerke
                   ) 

pROC_obj <- roc(SI_sWSR$diag,SI_sWSR$SI,
                direction= "<", #when cases have lower SI, we need to force the direction
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


auc_sWSR <- pROC_obj$auc

```

```{r save sWSR results}
write_tsv(sWSR_Set, paste0(dir, "sWSR.txt"))
write_tsv(SI_sWSR, file= paste0(dir, "SI_sWSR.txt"))
write.csv( unlist(SI_sWSR_ttest), file= paste0(dir,"SI_sWSR_ttest.txt"))
write.csv( unlist(SI_sWSR_wilcox), file= paste0(dir,"SI_sWSR_wilcox.txt"))
write.csv(r2l_sWSR,  file= paste0(dir,"r2l_sWSR.txt"))
write.csv(auc_sWSR[1], file=paste0(dir, "auc_sWSR.txt"))
ggsave(pROC, filename = paste0(dir, "ROC_sWSR.pdf"), width = 43, height = 43, units = "mm")  
ggsave(plot= SI_sWSR_boxplot, filename = paste0(dir, "SI_sWSR_boxplot.pdf"), width = 43, height = 43, units = "mm")
ggsave(plot = SI_sWSR_dens, filename = paste0(dir, "SI_density.pdf"), width = 43, height = 43, units = "mm")


```



```{r save result for combination}
sWSR_CMC_HBCC <- sWSR_Set
sWSR_CMC_Pitt <- sWSR_Set
sWSR_BipSeq_sACC <- sWSR_Set
sWSR_GVEX_SMRI <- sWSR_Set

```


```{r raster plot sWSR CMC-HBCC}
col2 <- colorRampPalette(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
                           "#4393C3", "#2166AC", "#053061"))

SI_sWSR_HBCC <- sWSR_CMC_HBCC %>% 
  group_by(SampleID) %>% 
  mutate(SI = sum(abs(sWSR))/n_distinct(gene_id)) %>% #use the mean weighted scaled residual value 
  ungroup() 


SI_C_BD <- SI_sWSR_HBCC  %>% select(SampleID, PrimaryDiagnosis, SI) %>%  distinct()


SI_C <- SI_C_BD %>% filter(PrimaryDiagnosis == "HC")
SI_BD <- SI_C_BD %>% filter(PrimaryDiagnosis == "BD")

SI_C <- SI_C %>% 
  group_by(SampleID) %>% 
  arrange(SI) %>% ungroup()

SI_C$SampleID <-
  factor(SI_C$SampleID, levels=unique(SI_C$SampleID))

SI_BD <- SI_BD %>% 
  group_by(SampleID) %>% 
  arrange(SI) %>% ungroup()

SI_BD$SampleID <-
  factor(SI_BD$SampleID, levels=unique(SI_BD$SampleID))




BD_SI_sWSR_HBCC  <- SI_sWSR_HBCC %>% filter(PrimaryDiagnosis == "BD") 
C_SI_sWSR_HBCC <- SI_sWSR_HBCC %>% filter(PrimaryDiagnosis == "HC")


BD_SI_sWSR_HBCC$SampleID <-
  factor(BD_SI_sWSR_HBCC$SampleID, levels=unique(SI_BD$SampleID))

BD_SI_sWSR_HBCC <-
  BD_SI_sWSR_HBCC %>% 
  group_by(gene_id) %>% 
  mutate(gene_mean=mean(abs(sWSR))) %>% 
  arrange(gene_mean) %>% ungroup() 

BD_SI_sWSR_HBCC$gene_id <-
  factor(BD_SI_sWSR_HBCC$gene_id, levels=unique(BD_SI_sWSR_HBCC$gene_id))


C_SI_sWSR_HBCC$SampleID <-
  factor(C_SI_sWSR_HBCC$SampleID, levels=unique(SI_C$SampleID))

C_SI_sWSR_HBCC <-
  C_SI_sWSR_HBCC %>% 
  group_by(gene_id) %>% 
  mutate(gene_mean=mean(abs(sWSR))) %>% 
  arrange(gene_mean) %>% ungroup() 

C_SI_sWSR_HBCC$gene_id <-
  factor(C_SI_sWSR_HBCC$gene_id, levels=unique(BD_SI_sWSR_HBCC$gene_id))





top_BD_sWSR_genes <- sWSR_CMC_HBCC %>% filter(PrimaryDiagnosis == "BD") %>% group_by(gene_id) %>%  summarise(mean_BD = mean(abs(sWSR))) %>% arrange(desc(mean_BD)) %>%  head(n=32) %>%  select(gene_id)

  BD_res <- 
    BD_SI_sWSR_HBCC %>% 
    filter(gene_id %in% top_BD_sWSR_genes$gene_id) %>% #filter out the 20 genes with lowest mean abs value in the BD
    ggplot(aes(x=SampleID, y=gene_id)) + 
  geom_raster(aes(fill=sWSR)) + 
      scale_fill_stepsn(colours = c("#2166AC", "#4393C3",  "#D1E5F0",   "#FFFFFF", "#FDDBC7", "#D6604D", "#B2182B"),
                        breaks = c(-2,-1,0, 1, 2), limits=c(-10,10)) +
           theme(axis.text= element_blank()) +
      theme(axis.ticks =  element_blank()) +
      theme(axis.title  =  element_blank()) +
      theme(legend.title = element_blank())+
    labs(title = "BD samples", ) +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Gene")

 C_res <- 
      C_SI_sWSR_HBCC %>% 
   filter(gene_id %in% top_BD_sWSR_genes$gene_id) %>% 
   ggplot(aes(x=SampleID, y=gene_id)) + 
  geom_raster(aes(fill=sWSR)) + 
      scale_fill_stepsn(colours = c("#2166AC", "#4393C3",  "#D1E5F0",   "#FFFFFF", "#FDDBC7", "#D6604D", "#B2182B"),
                        breaks = c(-2,-1,0, 1, 2), limits=c(-10,10)) +
    theme(axis.text.x= element_blank())+
      theme(axis.ticks.x =  element_blank()) +
      theme(axis.title.x  =  element_blank())+
   theme(legend.position = "none") +
    labs(title = "Control samples") +
    theme(plot.title = element_text(hjust = 0.5)) +
    ylab("Gene")  
 
#show the BD and C raster plot together with the tiles being the same size (number of samples drives the width of the plot)   
library(egg)
p1 <- egg::set_panel_size( C_res, width=unit(length(unique(C_SI_sWSR_HBCC$SampleID))/1.75, "mm"),
                          height=unit(length(subset)*1.5, "mm"), margin = unit(0.5, "mm") )
p2 <- egg::set_panel_size( BD_res, width=unit(length(unique(BD_SI_sWSR_HBCC$SampleID))/1.75, "mm"),
                          height=unit(length(subset)*1.5, "mm"), margin = unit(0.5, "mm") )


SI_C_plot <- 
   SI_C%>% mutate(gene_id= "SI_sample") %>% 
   ggplot(aes(x=SampleID, y=gene_id)) + 
     geom_raster(aes(fill=SI)) + 
     scale_fill_stepsn(colours = c("#FFFFFF",  "#D6604D", "#B2182B"),
           #            breaks = c(0, 0.15, 0.2, 0.28), limits=c(0,0.5))   +
                     #  breaks = c(0, 0.15, 0.2, 0.28), limits=c(0,0.5))   +
              breaks = c(0, 0.8, 1.2), limits=c(0,3))   +
   theme(axis.text.x= element_blank()) +
   theme(axis.text.y= element_blank()) +
      theme(axis.title.x = element_blank()) +
    theme(axis.ticks = element_blank()) +
   theme(legend.position = "none") +
    ylab("SI score")
 
p3 <- egg::set_panel_size( SI_C_plot, width=unit(length(unique(SI_C$SampleID))/1.75, "mm"),
                          height=unit(3, "mm"), margin = unit(0.1, 0.1,0.1,0.1, "mm") )



 SI_BD_plot <- 
   SI_BD %>% 
   mutate(gene_id= "SI_sample") %>% 
   ggplot(aes(x=SampleID, y=gene_id)) + 
     geom_raster(aes(fill=SI)) + 
     scale_fill_stepsn(colours = c("#FFFFFF",  "#D6604D", "#B2182B"),
                     #  breaks = c(0, 0.15, 0.2, 0.28), limits=c(0,0.5))   +
              breaks = c(0, 0.8, 1.2), limits=c(0,3))   +
   theme(axis.text.x= element_blank()) +
   theme(axis.text.y= element_blank()) +
   theme(axis.title = element_blank()) +
    theme(axis.ticks = element_blank())+
      theme(legend.title = element_blank()) 

 
p4 <- egg::set_panel_size( SI_BD_plot, width=unit(length(unique(SI_BD$SampleID))/1.75, "mm"),
                          height=unit(3, "mm"), margin = unit(0.1, 0.1,0.1,0.1, "mm") )



ggsave(plot =p1, filename = paste0(dir,"C_sWSR.pdf") )
ggsave(plot =p2, filename = paste0(dir,"BD_sWSR.pdf") )

ggsave(plot =p1, filename = paste0(dir,"C_sWSR.tiff") )
ggsave(plot =p2, filename = paste0(dir,"BD_sWSR.tiff") )

ggsave(plot =p1, filename = paste0(dir,"C_sWSR_thr2.tiff") )
ggsave(plot =p2, filename = paste0(dir,"BD_sWSR_thr2.tiff") )

ggsave(plot =p3, filename = paste0(dir,"C_sWSR_SI.pdf") )
ggsave(plot =p4, filename = paste0(dir,"BD_sWSR_SI.pdf") )
ggsave(plot =p3, filename = paste0(dir,"C_sWSR_SI.tiff") )
ggsave(plot =p4, filename = paste0(dir,"BD_sWSR_SI.tiff") )


```



When all 4 cortical sets are done - cross validation and predicted probabilities

```{r logistic regression based on sWSR sWSR split by dataset}
#import sWSR data

# sWSR_CMC_HBCC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/CMC_HBCC/sWSR/sWSR.txt")
# sWSR_CMC_Pitt <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/CMC-Pitt/sWSR/sWSR.txt")
# sWSR_GVEX_SMRI <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/GVEX-SMRI/sWSR/sWSR.txt")
# sWSR_BipSeq_sACC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/BipSeq-sACC/sWSR/sWSR.txt")
# 
# 
# sWSR_CMC_HBCC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff_lasso/CMC-HBCC/sWSR/sWSR.txt")
# sWSR_CMC_Pitt <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff_lasso/CMC-Pitt/sWSR/sWSR.txt")
# sWSR_GVEX_SMRI <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff_lasso/GVEX-SMRI/sWSR/sWSR.txt")
# sWSR_BipSeq_sACC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff_lasso/BipSeq-sACC/sWSR/sWSR.txt")
# 
# 
# 
# #"old" sWSR values, to test the new version of the code.
# # sWSR_CMC_HBCC <- read_tsv("/Users/asbjorh/PhD/RNAseq_temporal/analysis/OFC/01_CommonMind/06_lasso_model/RPKM/high_L2G_nearest/CM_str/wsr/wsr_t.txt")
# # sWSR_CMC_Pitt <- read_tsv("/Users/asbjorh/PhD/RNAseq_temporal/analysis/OFC/01_CommonMind/06_lasso_model/RPKM/high_L2G_nearest/CM_unstr/Pitt/wsr/wsr_t.txt")
# # sWSR_GVEX_SMRI <- read_tsv("/Users/asbjorh/PhD/RNAseq_temporal/analysis/OFC/01_BrainGVEX/06_lasso_model/results/RPKM/high_L2G_nearest/wsr/wsr_t.txt") 
# # sWSR_BipSeq_sACC <- read_tsv("/Users/asbjorh/PhD/RNAseq_temporal/analysis/OFC/01_BipSeq/06_lasso_model/RPKM/sACC/wsr/wsr_t.txt")
# # 
# 
# sWSR_CMC_HBCC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/original_gene_selection/CMC-HBCC/sWSR/sWSR.txt")
# sWSR_CMC_Pitt <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/original_gene_selection/CMC-Pitt/sWSR/sWSR.txt")
# sWSR_GVEX_SMRI <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/original_gene_selection/GVEX-SMRI/sWSR/sWSR.txt")
# sWSR_BipSeq_sACC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/original_gene_selection/BipSeq-sACC/sWSR/sWSR.txt")

#sWSR_CMC_HBCC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/CMC-HBCC/sWSR/sWSR.txt")
#sWSR_CMC_Pitt <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/CMC-Pitt/sWSR/sWSR.txt")
#sWSR_GVEX_SMRI <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/GVEX-SMRI/sWSR/sWSR.txt")
#sWSR_BipSeq_sACC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/BipSeq-sACC/sWSR/sWSR.txt")

```

```{r combine sWSR}
sWSR_all <- bind_rows("CMC-HBCC"= sWSR_CMC_HBCC,
                                 "CMC-Pitt" = sWSR_CMC_Pitt,
                                "GVEX-SMRI"= sWSR_GVEX_SMRI,  
                                "BipSeq-sACC"= sWSR_BipSeq_sACC, 
                                 .id = "Set")

sWSR_all <- sWSR_all %>% drop_na(sWSR)
sWSR_wide <- 
  sWSR_all %>% 
  select(Set, gene_id, SampleID, PrimaryDiagnosis, sWSR) %>% 
  pivot_wider(names_from = gene_id, values_from = sWSR)

sWSR_wide <- sWSR_wide %>%  select_if(~ !any(is.na(.))) # remove genes with missing values. EBLN3P is not in GVEX
sWSR_wide <- sWSR_wide %>% drop_na() #remove samples with missing values (probably REDUNDANT now)

sWSR_wide$PrimaryDiagnosis <- factor(sWSR_wide$PrimaryDiagnosis, levels = c("HC", "BD"))
```

```{r combine SI plots density and ROC}
#calculate SI
SI_sWSR_all <- sWSR_all %>% 
 drop_na() %>%  #when all models have R2=0 for a gene, then res will be 0 and SI NA
  group_by(SampleID) %>% 
  mutate(SI = sum(abs(sWSR))/n_distinct(gene_id)) %>% #use the mean weighted scaled residual value 
  ungroup() %>% 
  select(SampleID, PrimaryDiagnosis, SI, Set) %>%
  distinct()

SI_sWSR_all$Set <- factor(SI_sWSR_all$Set, levels = c("CMC-HBCC", "CMC-Pitt", "GVEX-SMRI", "BipSeq-sACC"))
SI_sWSR_all$PrimaryDiagnosis <- factor(SI_sWSR_all$PrimaryDiagnosis, levels = c("HC", "BD"))


SI_sWSR_all_dens <- 
  SI_sWSR_all %>% 
  ggplot(aes(x=SI, colour=PrimaryDiagnosis)) + 
  geom_density() +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC", "BD")) +
  theme(legend.position = "none",
        strip.background.x  = element_rect(fill  = "white"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()
        ) +
  labs(x= "SI score", y= "Density") +
    facet_wrap(~ Set, nrow = 1,scales = "free_x") +
    scale_x_continuous( breaks= c(0, 0.5, 1, 1.5), limits = c(0,2) ) 

#ggsave(SI_sWSR_all_dens, filename = "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/SI_dens_all_adjX.pdf", width = 9, height = 4.2, units= "cm")

ggsave(SI_sWSR_all_dens, filename = "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/fix_lambda/SI_dens_all_adjX.pdf", width = 9, height = 4.2, units= "cm")

pROC_obj <- list()

for(dataset in c("CMC-HBCC",    "CMC-Pitt" ,   "GVEX-SMRI",   "BipSeq-sACC" )){
  
  SI_sWSR <- SI_sWSR_all %>% filter(Set== dataset)
  
    # ROC calculations
  pROC_obj[[dataset]] <- pROC::roc(
    response= SI_sWSR$PrimaryDiagnosis, 
    predictor= SI_sWSR$SI, 
    levels= c("HC", "BD") ,
   direction= "<" #force the direction, i.e higher score for cases
)
}

# Applying the auc function to all the ROC curves
AUC_all <- sapply(pROC_obj, auc)
write.csv(AUC_all, paste0(dir, "AUC_all.txt"))


colours <- brewer.pal(n= 4, name = "Dark2")

# Plotting ROC curves
ROC_SI <- 
  ggroc(pROC_obj, aes= c("color", "alpha") )  +
  coord_cartesian(xlim= c(1,0), ylim= c(0,1)) +
  scale_color_manual(values= c(
    "CMC-HBCC" = "black", 
    "CMC-Pitt" = colours[1], 
    "GVEX-SMRI"= colours[2] ,  
    "BipSeq-sACC"= colours[3]
  )) +
  scale_alpha_manual(values= c(
    "CMC-HBCC" = 1, 
    "CMC-Pitt" = 0.5, 
    "GVEX-SMRI"= 0.5 ,  
    "BipSeq-sACC"= 0.5
  )) +
  theme(
    legend.title = element_blank(),
    legend.position = c(0.68, 0.22),
    # legend.position = "none",
    legend.direction = "vertical",
    legend.key.size = unit(2, "mm"),
    legend.background=element_rect(colour= "white"),
    legend.text = element_text(size=5),
    legend.margin = margin(0.2, 0.2, 0.2, 0.2, unit= "mm" )
  ) +
  geom_abline(slope= 1, intercept= 1, linetype= "dashed", alpha=0.2, colour= "black") +
  xlab("Specificity") +
  ylab("Sensitivity")


ggsave(plot =ROC_SI, filename = paste0(dir,"ROC_SI.pdf"), width = 43, height = 43, units = "mm" )



```




```{r logistic regression with cross validation}
#dir <- "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/CV_prob/"
#dir <- "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff/CV_prob/"
#dir <- "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_model/CV_prob/"
dir <- "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/CV_prob/"
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff_lasso/CV_prob/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/original_gene_selection/CV_prob/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/keep_TRPT1_LMAN2L/CV_prob/")


pROC_obj <- list()
glm_sWSR_4datasets <- list()
sWSR_wide_testset <- list()
pos_pred <- list() #positive predicted value

# Iterating over subsets to run logistic regression to predict diagnosis probabilities
# In each iteration, pick one set as test (predict probabilites) and the other 3 will be used to fit the logistic regression
for(testset in c("CMC-HBCC",    "CMC-Pitt" ,   "GVEX-SMRI",   "BipSeq-sACC" )){
  # Defining the test set
  sWSR_wide_test <- sWSR_wide[sWSR_wide$Set == testset,] # removing all rows that don't belong to the test set
  sWSR_wide_test <- sWSR_wide_test[, !colnames(sWSR_wide_test) %in% "Set"] #removing the Set column (no longer needed bcse working on one Set at a time)

  cat("\nTestset:")
  print(testset)
  cat("\n")
  
  #sanity check
cat("Dimensions of set used as test")
print(dim(sWSR_wide_test))


  # Defining the 3 modeling sets for logistic regression fitting
  sWSR_wide_set <- sWSR_wide[sWSR_wide$Set != testset,]
  sWSR_wide_set <- sWSR_wide_set[, !colnames(sWSR_wide_set) %in% "Set"] #remove Set column
  sWSR_wide_set <- sWSR_wide_set[, !colnames(sWSR_wide_set) %in% "SampleID"] #remove SampleID column

    #sanity check
cat("Dimensions of set used for modelling")
print(dim(sWSR_wide_set))

cat("\n total samples= ")
print(length(rownames(sWSR_wide_test)) + length(rownames(sWSR_wide_set)) )
   
  # Logistic regression
  glm_sWSR_set <- glm(PrimaryDiagnosis ~ . , data= sWSR_wide_set , family = "binomial")
  
  # Collecting the 4 logistic regressions.
  glm_sWSR_4datasets[[testset]] <- glm_sWSR_set
  
  # Get the predicted diag from the model
  probabilities <- predict.glm(
    object = glm_sWSR_set, 
    newdata = sWSR_wide_test, 
    type = "response")

cat("Mean correctly predicted class")
  predicted.classes <- ifelse(probabilities > 0.5, "BD", "HC")
  print(mean(predicted.classes == sWSR_wide_test$PrimaryDiagnosis))
  
  TP = length(predicted.classes[ predicted.classes == "BD" & predicted.classes == sWSR_wide_test$PrimaryDiagnosis])
  FP = length(predicted.classes[predicted.classes == "BD" & predicted.classes != sWSR_wide_test$PrimaryDiagnosis])

  pos_pred[[testset]] <- TP/(TP+FP)
  
  sWSR_wide_test$prob <- probabilities
  sWSR_wide_testset[[testset]] <- sWSR_wide_test
  
    # ROC calculations
  pROC_obj[[testset]] <- pROC::roc(
    response= sWSR_wide_test$PrimaryDiagnosis, 
    predictor= probabilities, 
    levels= c("HC", "BD") ,
   direction= "<" #force the direction, i.e higher score for cases
)
}

# Applying the auc function to all the ROC curves
AUC_all <- sapply(pROC_obj, auc)
write.csv(AUC_all, paste0(dir, "AUC_all.txt"))

# save positive predicted values
pos_pred_all <- do.call(rbind, pos_pred)
colnames(pos_pred_all) <- "PPV"

#save all probabilitites after regression with the 3 other cortical sets
sWSR_prob <- do.call(rbind, sWSR_wide_testset)


#add Set info to probabilities
sWSR_prob <- sWSR_prob %>% 
  select(SampleID, PrimaryDiagnosis, prob) %>% 
  left_join(select(info_all, SampleID, Set), by= "SampleID")

sWSR_prob$Set <- factor(sWSR_prob$Set, levels = c("CMC-HBCC", "CMC-Pitt","GVEX-SMRI", "BipSeq-sACC"))
```

```{r plot ROC curves}

colours <- brewer.pal(n= 4, name = "Dark2")

# Plotting ROC curves
ROC_glm_CV_sWSR <- 
  ggroc(pROC_obj, aes= c("color", "alpha") )  +
  coord_cartesian(xlim= c(1,0), ylim= c(0,1)) +
  scale_color_manual(values= c(
    "CMC-HBCC" = "black", 
    "CMC-Pitt" = colours[1], 
    "GVEX-SMRI"= colours[2] ,  
    "BipSeq-sACC"= colours[3]
  )) +
  scale_alpha_manual(values= c(
    "CMC-HBCC" = 1, 
    "CMC-Pitt" = 0.5, 
    "GVEX-SMRI"= 0.5 ,  
    "BipSeq-sACC"= 0.5
  )) +
  theme(
    legend.title = element_blank(),
    legend.position = c(0.68, 0.22),
    # legend.position = "none",
    legend.direction = "vertical",
    legend.key.size = unit(2, "mm"),
    legend.background=element_rect(colour= "white"),
    legend.text = element_text(size=5),
    legend.margin = margin(0.2, 0.2, 0.2, 0.2, unit= "mm" )
  ) +
  geom_abline(slope= 1, intercept= 1, linetype= "dashed", alpha=0.2, colour= "black") +
  xlab("Specificity") +
  ylab("Sensitivity")


ggsave(plot =ROC_glm_CV_sWSR, filename = paste0(dir,"ROC_glm_CV_sWSR.pdf"), width = 43, height = 43, units = "mm" )

```

```{r plot WSR boxplots}
plot_genes <- c("CACNA1C", "FADS2", "RASGRP1")

p <- 
  sWSR_all %>%
filter( gene_id %in% plot_genes) %>%
ggplot(aes(x = PrimaryDiagnosis, y = wsr, color = PrimaryDiagnosis)) +
  theme_bw() +
  geom_boxplot(outlier.shape = NA, colour="black") +
  geom_jitter(size=0.1,width = 0.2)  +
  facet_grid(factor(gene_id, levels= plot_genes) ~ factor(Set, levels = c("CMC-HBCC", "CMC-Pitt","GVEX-SMRI", "BipSeq-sACC")),
             scales = "free_y") +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC", "BD")) +
  scale_x_discrete(breaks=c(0,1), labels=c("HC", "BD")) +
  theme(text = element_text(size=5),
        legend.position="none",
        axis.ticks = element_blank(),
        strip.background.y  = element_rect(fill  = "#F4A582"),
         strip.background.x  = element_rect(fill  = "white")) + 
    xlab("") + ylab("") +
  coord_cartesian(ylim = c(-0.1, 0.12)) +
                    scale_y_continuous(
       labels = label_number(accuracy = 0.1)
  )

p_blue <- 
  p + theme(strip.background.y = element_rect(fill  = "#92C5DE")) + coord_cartesian(ylim = c(-0.2, 0.15)) 
p_white <- 
  p + theme(strip.background.y = element_rect(fill  = "white")) + coord_cartesian(ylim = c(-0.25, 0.25)) 


 ggsave(plot=p, path = dir, filename = "compare_WSR_BD_HC_redstrip.pdf", width = 9, height = 10.48, units = "cm")
  ggsave(plot=p_blue, path = dir, filename = "compare_WSR_BD_HC_bluestrip.pdf", width = 9, height = 10.48, units = "cm")
  ggsave(plot=p_white, path = dir, filename = "compare_WSR_BD_HC_whitestrip.pdf", width = 9, height = 10.48, units = "cm")

```




```{r investigating the predicted probabilities further}
sWSR_prob_wilcox <- sWSR_prob %>%  group_by(Set) %>% wilcox_test(prob ~ PrimaryDiagnosis, ref.group = "HC")


#get summary data from models in previous chunk
summary_all <- lapply(glm_sWSR_4datasets, summary)

#save the coefficients from all datasets, creating a column with set name
coeff_all <- lapply(seq_along(summary_all), #start the function for each element of list
                    function(i) summary_all[[i]]$coefficients %>%  #define function, "i" is undefined and becomes the number of the list elements currently worked on.
    as_tibble(rownames= "variable") %>%    
    arrange(desc(abs(Estimate))) %>%
    mutate(Set = names(summary_all)[[i]]))
 #combine into one df
coeff_all_df <- do.call(rbind, coeff_all)
coeff_all_wide <- coeff_all_df %>%   pivot_wider( names_from = Set, id_cols = variable, values_from = c( Estimate, `Std. Error`, `z value`,    `Pr(>|z|)`)) 

write.csv(coeff_all_df, paste0(dir, "coeff_all_df.txt"))
write_tsv(coeff_all_wide, paste0(dir, "coeff_all.txt"))

write.csv(pos_pred_all, paste0(dir, "PPV.txt"))

#visualize density of predicted probabilities by diagnosis
prob_glm_diag <- 
sWSR_prob %>%  
  ggplot(aes(x=prob, colour= PrimaryDiagnosis)) + 
  geom_density() +  
    facet_wrap(~Set, nrow = 1) + 
  scale_colour_manual(values=rev(cols)) +
  geom_vline(xintercept = 0.5, linetype= "dashed", alpha= 0.3)+ 
  scale_x_continuous(limits=c(0,1), breaks= c(0, 1) ) +
  scale_y_continuous( breaks = c(0, 1, 2), minor_breaks = seq(0, 2, 0.5))+
  theme(
    legend.position = "none",
    strip.background  = element_rect(fill  = "white"),
    strip.text = element_text(size=6, 
                              margin = margin(1,1,1,1, "mm")
    ),
    axis.text.y = element_blank(),
    axis.ticks.y =  element_blank() 
  ) +
  xlab("Predicted probabilities" )+
  ylab("Density")


#calculate the explanatory power of probability, r2l

r2l_prob <- list()
for(testset in c("CMC-HBCC",    "CMC-Pitt" ,   "GVEX-SMRI",   "BipSeq-sACC")){
  
  # logistic regression
  prob_glm <- glm(PrimaryDiagnosis ~ prob, data= sWSR_prob[sWSR_prob$Set == testset,], family = "binomial")
  pseudoR2 <- nagelkerke(prob_glm)
  
  
  if (pseudoR2$Pseudo.R.squared.for.model.vs.null[3] <= 0) { #skip the liability scaling if pseudoR2 is below zero
    next
  }
  
  
  frac_BD <- sWSR_prob %>%  filter(Set== testset) %>%  group_by(PrimaryDiagnosis) %>% summarise(n=n()) %>%  mutate(frac= n/sum(n)) %>% filter(PrimaryDiagnosis=="BD") %>% select(frac)
  
  r2l <- result_h2l(k=0.02, p=frac_BD$frac, r2n = pseudoR2$Pseudo.R.squared.for.model.vs.null[3] #take out the Nagelkerke
  )
  print(pseudoR2$Pseudo.R.squared.for.model.vs.null[3])
  r2l_prob[[testset]]      <- r2l
}

r2l_prob_all <- sapply(r2l_prob, cbind)
write.csv(r2l_prob_all, paste0(dir, "r2l_prob_all.txt"))
write_tsv(sWSR_prob_wilcox, paste0(dir, "sWSR_prob_wilcox.txt"))
write_tsv(sWSR_prob, paste0(dir, "sWSR_prob.txt"))
ggsave(prob_glm_diag, filename = paste0(dir, "prob_glm_diag.pdf"), width = 9.0, height = 4.2, units = "cm")

gene_list <- subset[subset %in% colnames(sWSR_wide)]
write.csv(gene_list, paste0(dir, "gene_list.txt"))

```


FInd genes where sWSR has predictive power, glm() on sWSR all cortical sets
```{r glm on sWSR all cortical sets}
sWSR_wide <- sWSR_wide %>% select(-Set, -SampleID)

glm_sWSR <- glm(PrimaryDiagnosis ~ . , data= sWSR_wide , family = "binomial")
summary_glm_sWSR_all <- summary(glm_sWSR)

glm_sWSR_all_coefficients <- summary_glm_sWSR_all$coefficients %>%  as_tibble(rownames= "variable") %>%    arrange(desc(abs(Estimate)))

write_tsv(glm_sWSR_all_coefficients, paste0(dir, "glm_sWSR_all_coefficients.txt"))

```



```{r plot sWSR DE}
order.CMC_HBCC.gene_id <- read.csv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/order.CMC_HBCC.gene_id.txt")
order.CMC_HBCC.gene_id <- order.CMC_HBCC.gene_id$x


stat.test_all<-
  sWSR_all %>% 
  #filter(wsr !=0) %>%  #wilcox test cannot handle many zero values well
  group_by(Set, gene_id) %>% 
  wilcox_test(data= ., formula = wsr ~ PrimaryDiagnosis, ref.group  = "HC" ) %>% 
  adjust_pvalue(method = "BH") %>% 
  mutate(y.position = 0.07) %>% 
  mutate(p.adj= round(p.adj, digits = 6))

stat.test_all$gene_id <- as.character(stat.test_all$gene_id)

#for Wilcoxon test, we need to calculate the direction in median difference 
p_thr <- 0.15#threshold for colour in tile
#p_thr <- 0.05 #only colour at least nominally significant

wsr_median <- 
  sWSR_all %>% 
    group_by(gene_id, Set, PrimaryDiagnosis) %>% 
    mutate(median= median(wsr)) %>% 
    ungroup() %>% 
  select(Set, gene_id, PrimaryDiagnosis, median) %>%  distinct() %>%
    group_by(gene_id, Set) %>% 
  pivot_wider( names_from = PrimaryDiagnosis, values_from = median) %>% 
  mutate(median_diff = BD- HC) %>% 
  left_join(select(stat.test_all, Set, gene_id, p, p.adj), by=c("Set", "gene_id")) %>% 
  mutate(direction_diff = case_when(
    median_diff > 0 & p < p_thr ~ "pos",
    median_diff < 0 & p < p_thr ~ "neg",
    TRUE ~ "no_diff"
  )) %>%  
  ungroup() %>%  distinct()


#combine data
stat.test_all <-
  stat.test_all %>% 
  left_join(select(wsr_median, gene_id, Set, median_diff, direction_diff), by= c("Set", "gene_id")) 

stat.test_all <- stat.test_all %>% drop_na() 



stat.test_all$Set <- factor(stat.test_all$Set, levels= c("CMC-HBCC","CMC-Pitt","GVEX-SMRI", "BipSeq-sACC"))

stat.test_all <- stat.test_all %>% mutate(sign = case_when(p.adj<= 0.05 ~ "FDR",
                                                           p <= 0.05 & p.adj>= 0.05 ~ "p",
                                                           TRUE ~ "none"))


stat.test_all$gene_id <- factor(stat.test_all$gene_id, levels=order.CMC_HBCC.gene_id)
stat.test_all <- stat.test_all %>% drop_na() 

statistic <- 
  ggplot(data=stat.test_all, aes(x=Set, y=gene_id)) + 
   geom_tile(aes(fill=direction_diff)) +   
         scale_fill_manual(values = c("pos" = "#F4A582", "neg" = "#92C5DE", "no_diff" = "#f7f7f7") #for wilcoxon 
                           ) +
     geom_point(data=stat.test_all, aes(size=sign, alpha=sign)) +
     scale_size_manual(values=c("FDR"=0.5, "p"=0.2, "none"= NA), guide="none") +
   scale_alpha_manual(values=c("FDR"=0.5, "p"=0.2, "none"= NA), guide="none") +
      theme(legend.title = element_blank())+
    labs(title = "" ) +
  theme_classic() +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(size = 5, vjust=1, angle= 60, hjust = 0.9),
          axis.text.y = element_text(size = 5, hjust=1),
          axis.ticks = element_blank(),
          legend.position = "none",
          axis.line = element_line(color='black', size = 0.2),
          panel.border = element_rect(colour = "black", fill=NA, size=0.2)
          ) + 
    scale_y_discrete() +
 scale_x_discrete(position="top") +
    theme(axis.text.x = element_text(hjust = 0.1, vjust= 1)) +
    ylab("") +
  xlab("")+
     geom_hline(yintercept=seq(1.5, length(unique(stat.test_all$gene_id))+1, 1), 
             lwd=0.1, colour="gray")# Horizonal lines

  
  
  
  ggsave(statistic, filename = paste0(dir, "sWSR_statistic_all_sets.pdf"),  width=4.9, height= 11.75, units="cm")
  ggsave(statistic, filename = paste0(dir, "sWSR_statistic_all_sets_pthr_0.15.pdf"),  width=4.9, height= 11.75, units="cm")
```


```{r polygenic architecture genes per ind}
thr <- 2

#count how many samples have disruption at +/- thr sd for genes
   SI2_number_thr<- sWSR_all  %>% 
    group_by(SampleID) %>% 
    mutate(sd_thr =  abs(sWSR) > thr) %>% add_tally(wt=sd_thr) %>%  
    select(SampleID,  PrimaryDiagnosis, n, Set) %>% distinct() 

   #insert median count by Set and diagnosis
  SI2_number_thr <- SI2_number_thr  %>% group_by(Set, PrimaryDiagnosis) %>%  mutate(median_n= median(n)) %>% ungroup() 
   
   SI2_number_GENEperIND_thr_combined <- 
SI2_number_thr  %>% group_by(Set, PrimaryDiagnosis) %>%  mutate(median_n= median(n)) %>% ungroup() %>% ggplot() + 
    geom_line(aes(x=n, color=Set, alpha=Set), stat="density") + 
      geom_vline(data=filter(SI2_number_thr, Set=="CMC-HBCC"), aes(xintercept = median_n), linetype = "dashed", alpha =0.3)+ #line with CMC-HBCC median
    facet_grid(cols =vars(PrimaryDiagnosis))+
    scale_color_manual(values= c(
        "CMC-HBCC" = "black", 
        "CMC-Pitt" = colours[1], 
        "GVEX-SMRI"= colours[2] ,  
        "BipSeq-sACC"= colours[3]
    )) +
    scale_alpha_manual(values= c(
        "CMC-HBCC" = 1, 
        "CMC-Pitt" = 0.5, 
        "GVEX-SMRI"= 0.5 ,  
        "BipSeq-sACC"= 0.5
    )) +
    theme(legend.position = "none",
          strip.background  = element_blank()
 
     #     strip.background  = element_rect(fill  = "white" )
          ) + 
    xlab("Number of imbalanced genes per individual") +
    ylab("Density (individuals)") +
    coord_cartesian(xlim = c(0, 15), ylim = c(0, 0.3))
   
    ggsave(SI2_number_GENEperIND_thr_combined , filename = paste0(dir, "GENEperIND.pdf") , height=4.3, width=8, units = "cm")
```


```{r polygenic architecture ind per gene}

sWSR_all <- bind_rows("CMC-HBCC"= sWSR_CMC_HBCC,
                                  "CMC-Pitt" = sWSR_CMC_Pitt,
                                 "GVEX-SMRI"= sWSR_GVEX_SMRI,  
                                 "BipSeq-sACC"= sWSR_BipSeq_sACC, 
                                  .id = "Set")
sWSR_samples <- sWSR_all %>% select(Set, SampleID, PrimaryDiagnosis) %>%  distinct() %>% group_by(Set, PrimaryDiagnosis) %>% mutate(n_ind= n()) %>% select(-SampleID) %>%   distinct() 


#SI_genes should be based on significance of coefficients in the logistic regression of all datasets
#We should see what the plot looks like with those 12 genes, but also perhaps consider setting the cutoff at 10% which would lead to 20 genes.

SI_genes <- glm_sWSR_all_coefficients %>%  filter(variable %in% subset, `Pr(>|z|)` < 0.05) %>% select(variable) %>% rename("gene_id" = variable)
#SI_genes <- glm_sWSR_all_coefficients %>%  filter(variable %in% subset, `Pr(>|z|)` < 0.10) %>% select(variable) %>% rename("gene_id" = variable)


#base top rank on significant coefficients in regression of 4 cortical sets
 sWSR_all <- sWSR_all %>% 
   mutate(SI_gene =
            ifelse(
              gene_id %in% SI_genes$gene_id, "top_rank",
              "Not_top_rank"
            ))


###extreme sWSR values: Individuals per gene

SI2_INDperGENE_thr_SI_genes <- 
  sWSR_all %>% 
  filter(SI_gene == "top_rank") %>% 
   left_join(sWSR_samples, by=c("Set", "PrimaryDiagnosis")) %>% #include total count of individuals per set and diagnosis
   group_by(Set, gene_id, PrimaryDiagnosis) %>% 
   mutate(sd_thr =  abs(sWSR) > thr) %>% #True if disrupted
   add_tally(wt=sd_thr) %>%  
    select(gene_id,  PrimaryDiagnosis, n, n_ind, Set, SI_gene) %>% distinct() %>% 
  mutate(n_norm = n/n_ind ) 

SI2_INDperGENE_thr_not_SI_genes <- 
  sWSR_all %>% 
  filter(SI_gene == "Not_top_rank") %>% 
   left_join(sWSR_samples, by=c("Set", "PrimaryDiagnosis")) %>% #include total count of individuals per set and diagnosis
   group_by(Set, gene_id, PrimaryDiagnosis) %>% 
   mutate(sd_thr =  abs(sWSR) > thr) %>% #True if disrupted
   add_tally(wt=sd_thr) %>%  
    select(gene_id,  PrimaryDiagnosis, n, n_ind, Set, SI_gene) %>% distinct() %>% 
  mutate(n_norm = n/n_ind ) 

SI2_INDperGENE_thr <- bind_rows(top_rank = SI2_INDperGENE_thr_SI_genes, Not_top_rank= SI2_INDperGENE_thr_not_SI_genes, .id= "SI_gene" )

 
 
SI2_prop_INDperGENE_thr_split <-
  SI2_INDperGENE_thr %>% 
  ggplot(aes(x=n_norm)) + 
      geom_line(data = filter(SI2_INDperGENE_thr, PrimaryDiagnosis== "HC"), aes(color=Set, alpha=Set), stat="density") + 
    geom_line(data = filter(SI2_INDperGENE_thr, PrimaryDiagnosis== "BD"),aes(color=Set, alpha=Set), stat="density") + 
 facet_grid(
    #rows=vars(PrimaryDiagnosis), 
             cols =  vars(factor(SI_gene, levels=c('top_rank', 'Not_top_rank'))),
           scales = "free_y") +
    scale_color_manual(values= c(
    "CMC-HBCC" = "black", 
    "CMC-Pitt" = colours[1], 
    "GVEX-SMRI"= colours[2] ,  
    "BipSeq-sACC"= colours[3], 
    "BipSeq-Amg"= colours[4]
  )) +
  scale_alpha_manual(values= c(
    "CMC-HBCC" = 1, 
    "CMC-Pitt" = 0.5, 
    "GVEX-SMRI"= 0.5 ,  
    "BipSeq-sACC"= 0.5, 
    "BipSeq-Amg"= 0.5
  )) +
  theme(legend.position = "none",
        #legend.position= c(0.8, 0.75),
        # strip.background.x  = element_rect(fill  = "white" )
        strip.background = element_blank( )
        ) + 
  xlab("Fraction of imbalanced individuals per gene") +
  ylab("Density (genes)")


#ggsave(SI2_prop_INDperGENE_thr_split, filename = paste0(dir, "INDperGENE_split_SI_gene_sign_coeff0.10.pdf"), width=8, height=4.3, units = "cm" )
ggsave(SI2_prop_INDperGENE_thr_split, filename = paste0(dir, "INDperGENE_split_SI_gene_sign_coeff0.05.pdf"), width=8, height=4.3, units = "cm" )

```


```{r rank genes based on BD wsr_t}

sum_sWSR_gene <- sWSR_all %>%
  filter(PrimaryDiagnosis == "BD") %>% 
  group_by(Set, gene_id) %>% 
  mutate(sum_sWSR = sum(abs(sWSR))) %>% 
  select(gene_id, Set, sum_sWSR) %>% 
  ungroup() %>% 
  distinct() %>%
  group_by(Set) %>% 
  arrange(desc(sum_sWSR)) %>% 
  ungroup()
  

#gene_order_t <- factor(sum_sWSR_gene$gene_id, levels = sum_sWSR_gene$gene_id)

```

```{r increasing genes AUC based on sWSR}

#dataset = "CMC-HBCC"

auc_GOI <- list()
gene_rank <- list()

for(dataset in c("CMC-HBCC", "CMC-Pitt", "GVEX-SMRI", "BipSeq-sACC") ){
  
  gene_order_t <- sum_sWSR_gene %>%  filter(Set== dataset) %>% select(gene_id)
  gene_order_t <- factor(gene_order_t$gene_id, levels = gene_order_t$gene_id)

SI_incr <- list()
SI_ttest_incr <- list()
SI_glm_incr <- list()
r2l_incr <- list()
pROCs_GOI_incr <- list()
  
for(genes in seq(from=1, to=length(gene_order_t), by= 1)){ 
  #nonsense with only 1 gene. From 2 to all modelled genes in the SI calculation
GOI <- gene_order_t[1:genes]
#calculate SI

SI <- sWSR_all %>% 
  filter(Set == dataset) %>% 
  filter(gene_id %in% GOI) %>% 
  group_by(SampleID) %>% 
  mutate(SI = sum(abs(sWSR))/n_distinct(gene_id)) %>% 
  ungroup()


SI$PrimaryDiagnosis <- as.factor(SI$PrimaryDiagnosis)

SI_incr[[genes]] <- SI

cat("genes included in SI score:")
print(length(GOI))



#SI_ttest_incr[[genes]] <- wilcox.test(SI ~ PrimaryDiagnosis, data = SI)
SI_glm_incr[[genes]] <- glm(PrimaryDiagnosis ~ SI, data= as.data.frame(SI), family = "binomial")


#get pseudo R2 from the regression
pseudoR2 <- nagelkerke(SI_glm_incr[[genes]])
 r2l_incr[[genes]] <- result_h2l(k=0.02, p=length(BD$SampleID)/(length(HC$SampleID) + length(BD$SampleID)), r2n = pseudoR2$Pseudo.R.squared.for.model.vs.null[3] #take out the Nagelkerke
                   ) 

pROC_obj <-  roc(SI$PrimaryDiagnosis,SI$SI,
                 levels = c("HC", "BD"),
               direction= "<", #should the cases have lower SI, we need to change the direction here
            smoothed = TRUE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)

pROCs_GOI_incr[[genes]] <- pROC_obj

}


auc_GOI_set <- pROCs_GOI_incr[1:length(gene_order_t)] %>% map_chr("auc")

auc_GOI_set <- as_tibble(auc_GOI_set) %>% 
  mutate("genes_in_SI"= seq(from=1, to=length(gene_order_t), by= 1)) %>% 
 rename("auc"= value)

auc_GOI_set$auc <- as.numeric(as.character(auc_GOI_set$auc))

#save the current dataset and gene names into list
auc_GOI[[dataset]] <- auc_GOI_set
gene_rank[[dataset]] <- gene_order_t
}

#combine all sets into one tibble
auc_GOI_all <- bind_rows(auc_GOI, .id= "Set" )


gene_rank_all <- 
  tibble::tibble(rank= gene_rank) %>% 
  unnest_wider(  rank, names_sep = "") %>% 
  mutate(Set = c("CMC-HBCC", "CMC-Pitt", "GVEX-SMRI", "BipSeq-sACC")) %>% 
  pivot_longer(cols = -Set, names_to = "rank", values_to = "gene_id")

auc_increasing_genes_sWSR_all  <- 
    auc_GOI_all %>%  
  ggplot(aes(x=genes_in_SI, y=auc, colour=Set, alpha=Set)) + 
  geom_point(size= 0.2) + 
  geom_line(aes(colour=Set, alpha= Set)) +
    scale_color_manual(values= c(
    "CMC-HBCC" = "black", 
    "CMC-Pitt" = colours[1], 
    "GVEX-SMRI"= colours[2] ,  
    "BipSeq-sACC"= colours[3]
  )) +
  scale_alpha_manual(values= c(
    "CMC-HBCC" = 1, 
    "CMC-Pitt" = 0.3, 
    "GVEX-SMRI"= 0.3 ,  
    "BipSeq-sACC"= 0.3
  )) +
  theme(
    legend.title = element_blank(),
   #legend.position = c(.4, .2),
   legend.position = "none",
   legend.direction = "vertical",
   legend.key.size = unit(2, "mm"),
   legend.background=element_blank(),
   legend.text = element_text()
    ) +
      geom_vline(xintercept = 15, linetype= "dashed", alpha= 0.3)+
  ylab("AUC") + xlab("Nb of genes used for SI score") +
  coord_cartesian(ylim= c(0.55, 0.80)) +
  scale_x_continuous(limits = c(0, 53), breaks = c(0, 10, 20, 30, 40, 50))


ggsave(plot =auc_increasing_genes_sWSR_all, filename = paste0(dir, "auc_increasing_genes_sWSR_all.pdf"), width = 43, height = 43, units = "mm" )

```

```{r SI score performance sWSR 15 top genes specific}
#the genes are ranked according to sWSR SI in BD in the list gene_rank

  setname <- list()
  ttest_p_15best <- list()
  AUC_15best <- list()  
  r2l_15best <- list()
  GOI_15best <- list()
  SI_15best <- list()
  
  
for(dataset in c("CMC-HBCC", "CMC-Pitt", "GVEX-SMRI", "BipSeq-sACC")){

#r SI score sWSR_15
#select n best genes within set
top_genes <- 10 
  
GOI <- gene_rank[[dataset]][1:top_genes]
  


  sWSR_all_set <- sWSR_all %>% 
    filter(Set == dataset, gene_id %in% GOI)
  
  
#calculate SI
SI_sWSR_15genes_best_set <- sWSR_all_set %>% 
    filter(gene_id %in% GOI) %>% 
    group_by(SampleID) %>% 
    mutate(SI = sum(abs(sWSR))/n_distinct(gene_id)) %>% #use the mean weighted scaled residual value 
    ungroup() %>%  
  select(SampleID, PrimaryDiagnosis,SI, Set) %>%  distinct() 

SI_sWSR_15genes_best_set$PrimaryDiagnosis <- factor(SI_sWSR_15genes_best_set$PrimaryDiagnosis, levels= c("HC", "BD"))
  
BD <- SI_sWSR_15genes_best_set %>% filter(PrimaryDiagnosis=="BD") %>%  select(SampleID) %>%  distinct()
HC <- SI_sWSR_15genes_best_set %>% filter(PrimaryDiagnosis=="HC") %>%  select(SampleID) %>%  distinct()

SI_sWSR_15genes_best_set_ttest <- wilcox.test(SI ~ PrimaryDiagnosis, data = SI_sWSR_15genes_best_set)

SI_sWSR_15genes_best_set_glm <- glm(PrimaryDiagnosis ~ SI, data= as.data.frame(SI_sWSR_15genes_best_set), family = "binomial")
summary_sWSR_15genes_best <- summary(SI_sWSR_15genes_best_set_glm)

#get pseudo R2 from the model
pseudoR2_sWSR_15genes_best <- nagelkerke(SI_sWSR_15genes_best_set_glm)
r2l_sWSR_15genes_best <- result_h2l(k=0.02, p=length(BD$SampleID)/(length(HC$SampleID) + length(BD$SampleID)), r2n = pseudoR2_sWSR_15genes_best$Pseudo.R.squared.for.model.vs.null[3] #take out the Nagelkerke
                   ) 

pROC_obj <- roc(SI_sWSR_15genes_best_set$PrimaryDiagnosis,SI_sWSR_15genes_best_set$SI,
                 levels = c("HC", "BD"),
               direction= "<", #should the cases have lower SI, we need to change the direction here
            smoothed = TRUE,
            # arguments for plot
            plot=FALSE)

AUC_15best[[dataset]] <- pROC_obj$auc
ttest_p_15best[[dataset]] <- SI_sWSR_15genes_best_set_ttest$p.value
r2l_15best[[dataset]] <- r2l_sWSR_15genes_best
setname[[dataset]] <- dataset
SI_15best[[dataset]] <- SI_sWSR_15genes_best_set
}

  results_15best_all <- bind_cols("Set" = unlist(setname), "w-test" = unlist(ttest_p_15best), "AUC"= unlist(AUC_15best), "R2"= unlist(r2l_15best))

  
 SI_all_15best <-  bind_rows("CMC-HBCC" = SI_15best$`CMC-HBCC`, "CMC-Pitt" = SI_15best$`CMC-Pitt`,  "GVEX-SMRI" = SI_15best$`GVEX-SMRI`, "BipSeq-sACC" =SI_15best$`BipSeq-sACC`, .id = "Set")
 


```

```{r save results 15 best within set}
#save_dir <-  "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/incr_gene_SI/15best/"

save_dir <-  "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/fix_lambda/incr_gene_SI/15best/"

write_tsv(gene_rank_all, file = paste0(save_dir, "gene_rank_all.txt" ))
write_tsv(SI_all_15best, file = paste0(save_dir, "SI_sWSR_15genes_best.txt" ))
write_tsv(results_15best_all, file= paste0(save_dir, "results_15best.txt"))

write_tsv(SI_all_15best, file = paste0(save_dir, "SI_sWSR_10genes_best.txt" ))
write_tsv(results_15best_all, file= paste0(save_dir, "results_10best.txt"))
```












```{r common rank based on BD sWSR}

rank_common_sWSR <- sWSR_all %>% 
  filter(PrimaryDiagnosis== "BD") %>% 
  group_by(gene_id) %>% 
  mutate(sWSR_mean_BD = mean(abs(sWSR)), na.rm= TRUE) %>% 
  select(gene_id, sWSR_mean_BD) %>%  distinct() %>% 
  ungroup() %>%  
  arrange(desc(sWSR_mean_BD)) %>% 
  mutate(rank = dense_rank(desc(sWSR_mean_BD)))



```

```{r increasing genes AUC based on sWSR common}
gene_order_t <- factor(rank_common_sWSR$gene_id, levels = rank_common_sWSR$gene_id)

auc_GOI <- list()
gene_rank <- list()

for(dataset in c("CMC-HBCC", "CMC-Pitt", "GVEX-SMRI", "BipSeq-sACC") ){

SI_incr <- list()
SI_ttest_incr <- list()
SI_glm_incr <- list()
r2l_incr <- list()
pROCs_GOI_incr <- list()
  
for(genes in seq(from=1, to=length(gene_order_t), by= 1)){ 
  #nonsense with only 1 gene. From 2 to all modelled genes in the SI calculation
GOI <- gene_order_t[1:genes]
#calculate SI

SI <- sWSR_all %>% 
  filter(Set == dataset) %>% 
  filter(gene_id %in% GOI) %>% 
  group_by(SampleID) %>% 
  mutate(SI = sum(abs(sWSR))/n_distinct(gene_id)) %>% 
  ungroup()


SI$PrimaryDiagnosis <- as.factor(SI$PrimaryDiagnosis)

SI_incr[[genes]] <- SI

cat("genes included in model:")
print(length(GOI))



#SI_ttest_incr[[genes]] <- wilcox.test(SI ~ PrimaryDiagnosis, data = SI)
SI_glm_incr[[genes]] <- glm(PrimaryDiagnosis ~ SI, data= as.data.frame(SI), family = "binomial")


#get pseudo R2 from the regression
pseudoR2 <- nagelkerke(SI_glm_incr[[genes]])
 r2l_incr[[genes]] <- result_h2l(k=0.02, p=length(BD$SampleID)/(length(HC$SampleID) + length(BD$SampleID)), r2n = pseudoR2$Pseudo.R.squared.for.model.vs.null[3] #take out the Nagelkerke
                   ) 

pROC_obj <-  roc(SI$PrimaryDiagnosis,SI$SI,
                 levels = c("HC", "BD"),
               direction= "<", #should the cases have lower SI, we need to change the direction here
            smoothed = TRUE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)

pROCs_GOI_incr[[genes]] <- pROC_obj

}


auc_GOI_set <- pROCs_GOI_incr[1:length(gene_order_t)] %>% map_chr("auc")

auc_GOI_set <- as_tibble(auc_GOI_set) %>% 
  mutate("genes_in_SI"= seq(from=1, to=length(gene_order_t), by= 1)) %>% 
 rename("auc"= value)

auc_GOI_set$auc <- as.numeric(as.character(auc_GOI_set$auc))

#save the current dataset and gene names into list
auc_GOI[[dataset]] <- auc_GOI_set
gene_rank[[dataset]] <- gene_order_t
}

#combine all sets into one tibble
auc_GOI_all <- bind_rows(auc_GOI, .id= "Set" )
gene_rank_all <- bind_rows(gene_rank, .id= "Set")

auc_increasing_genes_common_sWSR_all  <- 
    auc_GOI_all %>%  
  ggplot(aes(x=genes_in_SI, y=auc, colour=Set, alpha=Set)) + 
  geom_point(size= 0.2) + 
  geom_line(aes(colour=Set, alpha= Set)) +
    scale_color_manual(values= c(
    "CMC-HBCC" = "black", 
    "CMC-Pitt" = colours[1], 
    "GVEX-SMRI"= colours[2] ,  
    "BipSeq-sACC"= colours[3]
  )) +
  scale_alpha_manual(values= c(
    "CMC-HBCC" = 1, 
    "CMC-Pitt" = 0.3, 
    "GVEX-SMRI"= 0.3 ,  
    "BipSeq-sACC"= 0.3
  )) +
  theme(
    legend.title = element_blank(),
   #legend.position = c(.4, .2),
   legend.position = "none",
   legend.direction = "vertical",
   legend.key.size = unit(2, "mm"),
   legend.background=element_blank(),
   legend.text = element_text()
    ) +
      geom_vline(xintercept = 15, linetype= "dashed", alpha= 0.3)+
  ylab("AUC") + xlab("Nb of genes used for SI score") +
  coord_cartesian(ylim= c(0.55, 0.80)) +
  scale_x_continuous(limits = c(0, 53), breaks = c(0, 10, 20, 30, 40, 50))


ggsave(plot =auc_increasing_genes_common_sWSR_all, filename = paste0(dir, "auc_increasing_genes_common_rank_sWSR_all.pdf"), width = 43, height = 43, units = "mm" )

```





```{r sWSR_15_common}
#import sWSR
sWSR_all %>%  head()

#r SI score sWSR_15
#gene order:
#GOI <- rank_statistic_common$gene_id[1:15]
GOI <- rank_common_sWSR$gene_id[1:15]


#calculate SI
SI_sWSR_15genes_common <- sWSR_all %>% 
    filter(gene_id %in% GOI) %>% 
    group_by(SampleID, Set) %>% 
    mutate(SI = sum(abs(sWSR))/n_distinct(gene_id)) %>% #use the mean weighted scaled residual value 
    ungroup() %>%  
  select(SampleID, PrimaryDiagnosis, SI, Set) %>%  distinct() 

#SI_sWSR_15genes_common <- rename(SI_sWSR_15genes_common, "PrimaryDiagnosis" = PrimaryDiagnosis)

#keep only t-test, AUC and R2 liab. 

  setname <- list()
  ttest_p_15common <- list()
  AUC_15common <- list()  
  r2l_15common <- list()
  
for(set in c("CMC-HBCC", "CMC-Pitt", "GVEX-SMRI", "BipSeq-sACC")){

 
  SI_sWSR_15genes_common_set <- SI_sWSR_15genes_common %>% 
    filter(Set == set)
  
BD <- SI_sWSR_15genes_common_set %>% filter(PrimaryDiagnosis=="BD") %>%  select(SampleID) %>%  distinct()
HC <- SI_sWSR_15genes_common_set %>% filter(PrimaryDiagnosis=="HC") %>%  select(SampleID) %>%  distinct()

SI_sWSR_15genes_common_set_ttest <- wilcox.test(SI ~ PrimaryDiagnosis, data = SI_sWSR_15genes_common_set)

SI_sWSR_15genes_common_set_glm <- glm(PrimaryDiagnosis ~ SI, data= as.data.frame(SI_sWSR_15genes_common_set), family = "binomial")
summary_sWSR_15genes_common <- summary(SI_sWSR_15genes_common_set_glm)

#get pseudo R2 from the model
pseudoR2_sWSR_15genes_common <- nagelkerke(SI_sWSR_15genes_common_set_glm)
r2l_sWSR_15genes_common <- result_h2l(k=0.02, p=length(BD$SampleID)/(length(HC$SampleID) + length(BD$SampleID)), r2n = pseudoR2_sWSR_15genes_common$Pseudo.R.squared.for.model.vs.null[3] #take out the Nagelkerke
                   ) 

pROC_obj <- roc(SI_sWSR_15genes_common_set$PrimaryDiagnosis,SI_sWSR_15genes_common_set$SI,
                 levels = c("HC", "BD"),
               direction= "<", #should the cases have lower SI, we need to change the direction here
            smoothed = TRUE,
            # arguments for plot
            plot=FALSE)

AUC_15common[[set]] <- pROC_obj$auc
ttest_p_15common[[set]] <- SI_sWSR_15genes_common_set_ttest$p.value
r2l_15common[[set]] <- r2l_sWSR_15genes_common
setname[[set]] <- set
SI[[set]] <- set
}

  results_15common <- bind_cols("Set" = unlist(setname), "wilcox-test" = unlist(ttest_p_15common), "AUC"= unlist(AUC_15common), "R2"= unlist(r2l_15common))

```

```{r save results 15 common}


write_tsv(SI_sWSR_15genes_common, file = paste0(dir, "SI_sWSR_15genes_common.txt" ))
write_tsv(results_15common, file= paste0(dir, "results_15common.txt"))
```




```{r bar chart SI genes}
library(tidytext)
thr<- 2

#ordered bar-chart to count how many samples are disrupted at +/- thr sd for genes. Identifies which genes they are
#genes_ind_per_gene <- 

sWSR_all$Set <- factor(sWSR_all$Set, levels = c("CMC-HBCC", "CMC-Pitt","GVEX-SMRI", "BipSeq-sACC"))

genes_SI_BD <- 
  sWSR_all %>% 
     left_join(sWSR_samples, by=c("Set", "PrimaryDiagnosis")) %>% #include total count of individuals per set and diagnosis
   group_by(Set, gene_id, PrimaryDiagnosis) %>% 
   mutate(sd_thr =  abs(sWSR) > thr) %>% #True if disrupted
   add_tally(wt=sd_thr) %>%      
  select(gene_id,  PrimaryDiagnosis, n, n_ind, Set) %>% distinct() %>% 
  mutate(n_norm = n/n_ind ) %>% 
  filter(PrimaryDiagnosis== "BD") %>% 
   ggplot(aes(x=reorder_within(x = gene_id, by = -n, within= Set))) + 
    geom_col(aes(y=n_norm)) +
  scale_x_reordered()+
    theme(axis.text.x = element_text(angle=90, size = 6)) +
    xlab("gene_id") + 
    ylab(paste("Fraction BD samples with |sWSR| > 2")) +
    theme(axis.text.x = element_text(size=5, hjust=1, angle = 90, vjust = 1.1),
          strip.background.x  = element_rect(fill  = "white" )) + 
    facet_wrap(vars(factor(Set, levels = c("CMC-HBCC", "CMC-Pitt","GVEX-SMRI", "BipSeq-sACC",  "BipSeq-Amg"))), scales= "free",
               nrow= 3, ncol= 2)
 #facet_grid(cols=vars(Set), scales= "free")# +
 #facet_grid( rows=vars(factor(Set, levels = c("CMC-HBCC", "CMC-Pitt","GVEX-SMRI", "BipSeq-sACC",  "BipSeq-Amg"))), scales= "free")
#+ coord_flip()
  
ggsave(genes_SI_BD, filename = paste0(dir, "genes_prop_thr_BD.pdf"), width= 18,  units= "cm")


```


control experiment with HC vs HC



```{r control experiment HC- HC analysis}
dataset <- "CMC-HBCC"
dataset <- "BipSeq-sACC"

#split HBCC, BipSeq HC in 2, and run modelling and SI score analysis
set.seed(22)
#HBCC
HC1 <- info_CM %>% filter(`PrimaryDiagnosis`=="HC", Study=="CMC_HBCC", Age>=18) %>% filter(!SampleID %in% remove_PCA) %>%
  select(SampleID) %>% slice_sample(prop = 0.5, replace = FALSE)
HC2 <- info_CM %>% filter(`PrimaryDiagnosis`=="HC", Study=="CMC_HBCC", Age>=18) %>% filter(!SampleID %in% remove_PCA) %>%
  select(SampleID)  %>% filter(!SampleID %in% HC1$SampleID)

#rpkm_log <- rpkm_CMC
rpkm_log <- rpkm_log_CMC_HBCC

rpkm_log <- rpkm_log_BipSeq_sACC
info <- info_all %>%  filter(Set== dataset) 


HC <- info %>% filter(PrimaryDiagnosis=="HC") %>% select(SampleID)
set.seed(22) #1st round
#set.seed(222) #2nd round
#BipSeq sACC

HC1 <- HC %>%  slice_sample(prop = 0.6, replace = FALSE) #more samples to the controls, since we have fewer in BipSeq than HBCC
HC2 <- HC %>%  filter(!SampleID %in% HC1$SampleID)





#Which variables (covarites) should be fixed?
#for CMC, GVEX where ethnicity differs 
   p.fac = rep(1, length(subset) -1 + 5) 
    p.fac[1:5] = 0


#for BipSeq where ethnicity does not differ
    p.fac = rep(1, length(subset) -1 + 4) 
    p.fac[1:4] = 0

```

```{r run iterations of lasso test HCvsHC}
debug = TRUE

subset <- subset[subset %in% rownames(rpkm_log)]
subset <- unique(subset) 

#RNAseq input is rpkm_log
#SampleID are defined in
#HC$SampleID 
#HC2$SampleID
#info

it <- 200


model_fraction <- length(HC1$SampleID) *0.9 

# #2nd run: swap HC1 and HC2 
# HC1_2 <- HC2
# HC2_2 <- HC1
# HC1 <- HC1_2
# HC2 <- HC2_2

#save the results from iterations to list
res_std_HC1_iterate <- list()
res_std_HC2_iterate<- list()
R2_iterate <- list()  
models_iterate <- list()  
  

  #randomly picked sets of controls for the modeling
for(i in 1:it){
  set.seed(i) 
  HC1_model <- slice_sample(HC1, n= model_fraction) 
  modelset <- rpkm_log[rownames(rpkm_log) %in% subset,colnames(rpkm_log) %in% HC1_model$SampleID]
  testset1  <-  rpkm_log[rownames(rpkm_log) %in% subset,colnames(rpkm_log) %in% HC1$SampleID & !colnames(rpkm_log) %in% HC1_model$SampleID] 
  testset2 <- rpkm_log[rownames(rpkm_log) %in% subset,colnames(rpkm_log) %in% HC2$SampleID]
  lasso_models <- lasso_model(modelset = modelset, GOI = subset, info =info) #these functions are defined in the chunk "multiple linear model all GOI functions"
  R2 <- sapply(lasso_models, function(x) as.vector(x[["dev.ratio"]]))
  test_HC1 <- lasso_test(testset = testset1, models  = lasso_models , GOI = subset, info =info )
  test_HC2 <- lasso_test(testset = testset2, models = lasso_models , GOI = subset , info =info)
  models_iterate[[i]] <- lasso_models 
  res_std_HC1_iterate[[i]] = test_HC1 
  res_std_HC2_iterate[[i]] =  test_HC2
  R2_iterate[[i]] = R2 #save the R2 from all models from the iteration
    #needs sanity check. Is the length of rownames(test_HC1) == rownames(test_HC2) == R2_iterate[[i]]. Stop if not "all genes are not modelled. check gene subset for duplicates".
  # Modulus operation
  if(i %% 10==0) {
      # Print on the screen message
      cat(paste0("iteration: ", i, "\n"))
  }
  
  if(debug==TRUE & i == 10) {
      cat("HC1 res \n")
      print(matrix(test_HC1[1:4,1:4], nrow= 4, ncol=4, byrow = FALSE)) 
      print(colnames(test_HC1)[1:4])
      print(rownames(test_HC1)[1:4])
      cat("\nHC1 res has dimensions: \n")
      print(dim(test_HC1))
      cat("\nHC2 res \n")
      print(matrix(test_HC2[1:4,1:4], nrow= 4, ncol=4, byrow = FALSE))
      print(colnames(test_HC2)[1:4])
      print(rownames(test_HC2)[1:4])
      cat("\nHC2 res has dimensions: \n")
      print(dim(test_HC2))
      cat("\nR2 \n")
      cat(paste0(R2[1:4]), "\n")
      print(names(R2)[1:4])
      cat("number of R2:")
      print(length(R2))
  }
}


```



```{r calculate SI from residuals HC vs HC}

#import residuals and R2 if needed
#res_std_HC1_iterate
#res_std_HC2_iterate
#R2_iterate 

#weight residuals with R2
res_w_HC1_iterate <- lapply(1:length(res_std_HC1_iterate), function(x) res_std_HC1_iterate[[x]]*R2_iterate[[x]] ) #weight residuals according to R2 from the models: multiplication by the vector (R2) is by default done column-wise, since genes are in rows.
res_w_HC2_iterate <- lapply(1:length(res_std_HC2_iterate), function(x) res_std_HC2_iterate[[x]]*R2_iterate[[x]] )


#melt function within list
meltfun <- function(x,it) {
   z <- reshape2::melt(x,id.vars=rownames(x))
  colnames(z) <- c('gene_id', 'SampleID', 'res')
  z$it = it
   return(z)
}


res_w_HC1_all_long <- mapply(meltfun,res_w_HC1_iterate, it= seq(1:length(res_w_HC1_iterate)),SIMPLIFY=FALSE) #melt each iteration to long
res_w_HC1_all<-  do.call("rbind", res_w_HC1_all_long) #combine all iterations into one dataframe

res_w_HC2_all_long <- mapply(meltfun,res_w_HC2_iterate, it= seq(1:length(res_w_HC2_iterate)),SIMPLIFY=FALSE) #melt each iteration to long
res_w_HC2_all<-  do.call("rbind", res_w_HC2_all_long) #combine all iterations into one dataframe

#combine HC1 and HC2
res_w_all <- rbind(res_w_HC1_all, res_w_HC2_all )

#calculate mean weighted residual across iterations
mean_res_w <- res_w_all %>%  
  group_by(gene_id, SampleID) %>%  
  summarise(mean_res_w= mean(res,  na.rm=TRUE)) %>% 
  ungroup()




#calculate SI
SI <- mean_res_w %>% 
  group_by(SampleID) %>% 
  summarise(SI = sum(abs(mean_res_w))/n_distinct(gene_id)) %>% #use the mean weighted scaled residual value 
  ungroup()
  
SI <- SI %>% left_join(select(info, SampleID, PrimaryDiagnosis), by= "SampleID")


SI <- rename(SI, "diag" = PrimaryDiagnosis)



```




```{r compare C and BD HC vs HC}

 options(digits = 5)
SI_HC1 <- filter(SI, SampleID %in% HC1$SampleID) 
SI_HC2 <- filter(SI,  SampleID %in% HC2$SampleID)
ks.test( SI_HC2$SI,  SI_HC1$SI)
kstest <- ks.test( SI_HC2$SI,  SI_HC1$SI)


```


```{r plot SI density HC vs HC }
#dir <-  "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/HCvsHC/CMC-HBCC/"
dir <-  "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/HCvsHC/BipSeq-sACC/"

SI <- SI %>% 
  mutate(split_set = ifelse(SampleID %in% HC1$SampleID, 
                            "HC1", "HC2")
         )


 #show the distribution of SI values as points
SI_HCvsHC_boxplot <- 
  SI %>% 
 ggplot( aes(y=SI, x=split_set, colour= split_set))+ 
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(size=0.2) +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC1", "HC2")) +
  labs(colour="split_set") + 
  xlab("split_set")+ 
  theme(legend.key.size = unit(0.5,"line"),
        legend.position = ""
        ) + 
  scale_x_discrete(breaks=c(0,1),
                    labels=c("HC1", "HC2")) +
  ylab("SI score")

SI_HCvsHC_wtest <- wilcox.test(SI ~ split_set, data = SI) 

#is the variance equal?
#SI_HCvsHC_levene_test_variance <-  
 # leveneTest(SI ~ split_set, data = SI)

#dir <- "/Users/asbjorh/PhD/RNAseq_temporal/analysis/OFC/01_CommonMind/06_lasso_model/RPKM/high_L2G_nearest/CM_str/SI_after_iterations/SI_mean_wsr/HCvsHC/"

#dir <- "/Users/asbjorh/PhD/RNAseq_temporal/analysis/OFC/01_CommonMind/06_lasso_model/RPKM/high_L2G_nearest/CM_str/SI_after_iterations/SI_mean_wsr/HCvsHC/swap_HC_split/"

#dir <- "/Users/asbjorh/PhD/RNAseq_temporal/analysis/OFC/01_BipSeq/06_lasso_model/RPKM/sACC/SI_after_iterations/SI_mean_wsr/HCvsHC/"


write_tsv(SI, file = paste0(dir,"SI_wsr.txt"))
write_tsv(mean_res_w, paste0(dir, "wsr.txt"))
write.csv( unlist(SI_HCvsHC_wtest), file=paste0(dir, "SI_wtest.txt"))
write.csv(as.data.frame(unlist(kstest)), file = paste0(dir,"SI_kstest.txt"))
#write.csv(as.data.frame(unlist(SI_HCvsHC_levene_test_variance)), file = paste0(dir,"SI_levene_variance.txt"))
ggsave(plot= SI_HCvsHC_boxplot, filename = paste0(dir, "SI_boxplot_HCvsHC.pdf"), width = 43, height = 43, units = "mm")


```


```{r sWSR analysis HCvsHC}
wsr_HCvsHC <- mean_res_w %>% left_join(select(SI, SampleID, split_set), by ="SampleID")
wsr_HCvsHC <- rename(wsr_HCvsHC, "wsr" = mean_res_w)


#calculate mean and sd for the HC
wsr_mean_HC1 <- wsr_HCvsHC %>% 
  filter(split_set =="HC1") %>% 
  group_by(gene_id) %>% 
  mutate(mean_HC1 = mean(wsr),
         sd_HC1 = sd(wsr)) %>% 
  ungroup() %>% 
  select(gene_id, mean_HC1, sd_HC1) %>% 
  distinct()

#calculate sWSR This is how many sd(HC) are the individual value from the mean HC1?
#
wsr_HCvsHC <- wsr_HCvsHC %>% 
  left_join(wsr_mean_HC1, by="gene_id") %>% 
  mutate(sWSR = (wsr- mean_HC1)/ sd_HC1)


#calculate SI
SI_sWSR <- wsr_HCvsHC %>%
  filter(sd_HC1!= 0) %>% 
  group_by(SampleID) %>% 
  mutate(SI = sum(abs(sWSR))/n_distinct(gene_id)) %>% #use the mean weighted scaled residual value 
  ungroup() %>% 
  select(SampleID, split_set, SI) %>% 
  distinct()


#calculate ttest sWSR HCvsHC}
SI_sWSR_wtest_HCvsHC <- wilcox.test(SI ~ split_set, data = SI_sWSR)



write_tsv(SI_sWSR, file = paste0(dir,"SI_sWSR.txt"))
write.csv( unlist(SI_sWSR_wtest_HCvsHC), file=paste0(dir, "SI_sWSR_ttest.txt"))
```



```{r plot wsr statistics HCvsHC}

#import WSR data
WSR_CMC_HBCC <- read_tsv( "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/HCvsHC/CMC-HBCC/wsr.txt")

#WSR_CMC_HBCC <- wsr_HCvsHC

WSR_BipSeq_sACC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/HCvsHC/BipSeq-sACC/wsr.txt")

#WSR_BipSeq_sACC <- wsr_HCvsHC

wsr_HCvsHC_all <- bind_rows("CMC-HBCC"= WSR_CMC_HBCC,
                                "BipSeq-sACC"= WSR_BipSeq_sACC, 
                            
                                .id = "Set")

wsr_HCvsHC_all <- wsr_HCvsHC_all %>%  rename("wsr"= mean_res_w)
#import SI data
SI_CMC_HBCC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/HCvsHC/CMC-HBCC/SI_wsr.txt")
SI_BipSeq_sACC <- read_tsv("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/HCvsHC/BipSeq-sACC/SI_wsr.txt")


SI_wsr_HCvsHC_all <- bind_rows("CMC-HBCC"= SI_CMC_HBCC,
                                "BipSeq-sACC"= SI_BipSeq_sACC, 
                                .id = "Set")

wsr_HCvsHC_all <- wsr_HCvsHC_all %>% 
  left_join(select(SI_wsr_HCvsHC_all, SampleID, SI, split_set), by= "SampleID")

stat.test_all_HCvsHC <- wsr_HCvsHC_all %>% 
  filter(wsr !=0) %>%  #wilcox test cannot handle zero values well
  group_by(Set, gene_id) %>% 
  wilcox_test(data= ., formula = wsr ~ split_set, ref.group  = "HC1" ) %>% 
  #t_test(wsr ~ split_set, ref.group  = "HC1" ) %>% 
  adjust_pvalue(method = "BH") %>% 
  mutate(y.position = 0.07) %>% 
  mutate(p.adj= round(p.adj, digits = 6))




#p_thr <- 0.32 #threshold for colour in tile
p_thr <- 0.15
 
#for Wilcoxon test, we need to calculate the direction in median difference 
wsr_median <- 
  wsr_HCvsHC_all %>% 
    group_by(gene_id, Set, split_set) %>% 
    mutate(median= median(wsr)) %>% 
    ungroup() %>% 
  select(Set, gene_id, split_set, median) %>%  distinct() %>% 
    group_by(gene_id, Set) %>% 
  pivot_wider( names_from = split_set, values_from = median) %>% 
    left_join(select(stat.test_all_HCvsHC, Set, gene_id, p, p.adj), by=c("Set", "gene_id")) %>% 
  mutate(median_diff = HC2- HC1) %>% 
  mutate(direction_diff = case_when(
    median_diff > 0 & p < p_thr ~ "pos",
    median_diff < 0 & p < p_thr ~ "neg",
    TRUE ~ "no_diff"
  )) %>%  
  ungroup() %>%  distinct()


#combine data
stat.test_all_HCvsHC <-
  stat.test_all_HCvsHC %>% 
  left_join(select(wsr_median, gene_id, Set, median_diff, direction_diff), by= c("Set", "gene_id")) 

stat.test_all_HCvsHC$gene_id <- factor(stat.test_all_HCvsHC$gene_id, levels=order.CMC_HBCC.gene_id)
stat.test_all_HCvsHC$Set <- factor(stat.test_all_HCvsHC$Set, levels= c("CMC-HBCC", "BipSeq-sACC"))

stat.test_all_HCvsHC <- stat.test_all_HCvsHC %>% mutate(sign = case_when(p.adj<= 0.05 ~ "FDR",
                                                           p <= 0.05 & p.adj>= 0.05 ~ "p",
                                                           TRUE ~ "none"))

stat.test_all_HCvsHC <- stat.test_all_HCvsHC %>% drop_na() 

#

statistic <- 
  ggplot(data=stat.test_all_HCvsHC, aes(x=Set, y=gene_id)) + 
 # geom_tile(aes(fill=statistic)) + #for ttest
   geom_tile(aes(fill=direction_diff)) +   #for wilcoxon
         scale_fill_manual(values = c("pos" = "#F4A582", "neg" = "#92C5DE", "no_diff" = "#f7f7f7") #for wilcoxon
                           ) + 
#      scale_fill_stepsn(colours = c( "#F4A582", "#f7f7f7", "#92C5DE" ),     #for ttest
  #                          breaks = c(-1, 1), limits=c(-5,5)) +
     geom_point(aes(size=sign, alpha=sign)) +
     scale_size_manual(values=c("FDR"=0.5, "p"=0.2, "none"= NA), guide="none") +
   scale_alpha_manual(values=c("FDR"=0.5, "p"=0.2, "none"= NA), guide="none") +
      theme(legend.title = element_blank())+
    labs(title = "" ) +
  theme_classic() +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(size = 5, vjust=1, angle= 45, hjust = 0.9),
          axis.text.y = element_text(size = 5, hjust=1),
          axis.ticks = element_blank(),
          legend.position = "none",
          axis.line = element_line(color='black', size = 0.2),
          panel.border = element_rect(colour = "black", fill=NA, size=0.2)
          ) + 
 # scale_x_discrete(position="top") +
    ylab("") +
  xlab("")+
     geom_hline(yintercept=seq(1.5, length(unique(stat.test_all_HCvsHC$gene_id))+1, 1), 
             lwd=0.1, colour="gray")# Horizonal lines


ggsave(statistic, filename = "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/HCvsHC/statistic_all_sets_HCvsHC.pdf", width=3.32, height= 11.75, units="cm")


plot_genes <- c("CACNA1C", "FADS2", "RASGRP1")

p <- wsr_HCvsHC_all %>% 
filter( gene_id %in% plot_genes) %>%
ggplot(aes(x = split_set, y = wsr, color = split_set)) +
  theme_bw() +
  geom_boxplot(outlier.shape = NA, colour="black") +
  geom_jitter(size=0.1,width = 0.2)  +
#facet_wrap(~ factor(gene_id, levels=genes_R2$gene_id), nrow  = 1) +
  facet_grid(factor(gene_id, levels= plot_genes) ~ factor(Set, levels = c("CMC-HBCC",  "BipSeq-sACC",  "BipSeq-Amg"))) +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC1", "HC2")) +
#  scale_x_discrete(breaks=c(0,1), labels=c("HC1", "HC2")) +
  theme(text = element_text(size=5),
        legend.position="none",
        axis.ticks = element_blank(),
        strip.background.y  = element_rect(fill  = "white"),
         strip.background.x  = element_rect(fill  = "white"))  + ylab("") +
  coord_cartesian(ylim = c(-0.15, 0.15)) 

ggsave(plot=p, path = "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/HCvsHC/", filename = "compare_wsr_HCvsHC.pdf", width = 5.4, height = 10.48, units = "cm")

```



negative control base model on BD as modelset

```{r prepare dataset BD model}


dataset <- "BipSeq-sACC"
dataset <- "CMC-HBCC"
dataset <- "GVEX-SMRI"

info <- info_all %>%  filter(Set== dataset) 
#rpkm_log <- rpkm_log_all[,colnames(rpkm_log_all) %in% info$SampleID]
#rpkm_log <- rpkm_log_GVEX_SMRI

  
rpkm_log <- rpkm_log_CMC_HBCC 


# rpkm_log <- rpkm_log_BipSeq_sACC
 rpkm_log <- rpkm_log_GVEX_SMRI

subset <- subset[subset %in% rownames(rpkm_log)]

HC <- info %>% filter(PrimaryDiagnosis=="HC") %>% select(SampleID)
BD <-  info %>% filter(PrimaryDiagnosis=="BD") %>% select(SampleID)

setequal(info$SampleID, colnames(rpkm_log))


#Which variables (covarites) should be fixed?
#for CMC, GVEX where ethnicity differs 
   p.fac = rep(1, length(subset) -1 + 5) 
    p.fac[1:5] = 0


#for BipSeq where ethnicity does not differ
    p.fac = rep(1, length(subset) -1 + 4) 
    p.fac[1:4] = 0
```



```{r model with BD}


debug = TRUE

subset <- subset[subset %in% rownames(rpkm_log)]
subset <- unique(subset) 

it <- 100 #number of iterations  
#model_fraction <- length(BD$SampleID) *2/3 #for BipSeq sACC
model_fraction <- length(BD$SampleID) *0.9 #for CMC-HBCC, GVEX-SMRI

#save the results from iterations to list
res_std_HC_iterate <- list()
res_std_BD_iterate<- list()
R2_iterate <- list()  
models_iterate <- list()  
  

  #randomly picked sets of controls for the modeling
for(i in 1:it){
  set.seed(i) 
  BD_model <- slice_sample(BD, n= model_fraction) 
  modelset <- rpkm_log[rownames(rpkm_log) %in% subset,colnames(rpkm_log) %in% BD_model$SampleID]
  testset1 <- rpkm_log[rownames(rpkm_log) %in% subset,colnames(rpkm_log) %in% BD$SampleID & !colnames(rpkm_log) %in% BD_model$SampleID]
  testset2 <- rpkm_log[rownames(rpkm_log) %in% subset,colnames(rpkm_log) %in% HC$SampleID]
  lasso_models <- lasso_model(modelset = modelset, GOI = subset, info =info) #these functions are defined in the chunk "multiple linear model all GOI functions"
  R2 <- sapply(lasso_models, function(x) as.vector(x[["dev.ratio"]]))
  test_BD <- lasso_test(testset = testset1, models  = lasso_models , GOI = subset, info =info )
  test_HC <- lasso_test(testset = testset2, models = lasso_models , GOI = subset , info =info)
  models_iterate[[i]] <- lasso_models 
  res_std_BD_iterate[[i]] = test_BD 
  res_std_HC_iterate[[i]] =  test_HC
  R2_iterate[[i]] = R2 #save the R2 from all models from the iteration

  # Modulus operation
  if(i %% 10==0) {
      # Print on the screen some message
      cat(paste0("iteration: ", i, "\n"))
  }
  
  if(debug==TRUE & i == 10) {
      cat("BD res \n")
      print(matrix(test_BD[1:4,1:4], nrow= 4, ncol=4, byrow = FALSE)) 
      print(colnames(test_BD)[1:4])
      print(rownames(test_BD)[1:4])
      cat("\nBD res has dimensions: \n")
      print(dim(test_BD))
      cat("\nHC res \n")
      print(matrix(test_HC[1:4,1:4], nrow= 4, ncol=4, byrow = FALSE))
      print(colnames(test_HC)[1:4])
      print(rownames(test_HC)[1:4])
      cat("\nHC res has dimensions: \n")
      print(dim(test_HC))
      cat("\nR2 \n")
      cat(paste0(R2[1:4]), "\n")
      print(names(R2)[1:4])
      cat("number of R2:")
      print(length(R2))
  }
}

#weight residuals with R2
res_w_HC_iterate <- lapply(1:length(res_std_HC_iterate), function(x) res_std_HC_iterate[[x]]*R2_iterate[[x]] ) #weight residuals according to R2 from the models: multiplication by the vector (R2) is by default done column-wise, since genes are in rows.
res_w_BD_iterate <- lapply(1:length(res_std_BD_iterate), function(x) res_std_BD_iterate[[x]]*R2_iterate[[x]] )


#melt function within list
meltfun <- function(x,it) {
   z <- reshape2::melt(x,id.vars=rownames(x))
  colnames(z) <- c('gene_id', 'SampleID', 'res')
  z$it = it
   return(z)
}


res_w_HC_all_long <- mapply(meltfun,res_w_HC_iterate, it= seq(1:length(res_w_HC_iterate)),SIMPLIFY=FALSE) #melt each iteration to long
res_w_HC_all<-  do.call("rbind", res_w_HC_all_long) #combine all iterations into one dataframe

res_w_BD_all_long <- mapply(meltfun,res_w_BD_iterate, it= seq(1:length(res_w_BD_iterate)),SIMPLIFY=FALSE) #melt each iteration to long
res_w_BD_all<-  do.call("rbind", res_w_BD_all_long) #combine all iterations into one dataframe

#combine C and BD
res_w_all <- rbind(res_w_HC_all, res_w_BD_all )

#calculate mean weighted residual across iterations
mean_res_w <- res_w_all %>%  
  group_by(gene_id, SampleID) %>%  
  summarise(mean_res_w= mean(res,  na.rm=TRUE)) %>% 
  ungroup()




#calculate SI
SI <- mean_res_w %>% 
  group_by(SampleID) %>% 
  summarise(SI = sqrt(sum(mean_res_w^2)/n_distinct(gene_id))) %>% 
  ungroup()
  
SI <- SI %>% left_join(select(info, SampleID, PrimaryDiagnosis), by= "SampleID")


SI <- rename(SI, "diag" = PrimaryDiagnosis)



#save the residuals and SI-scores
write_tsv(SI, file = paste0(dir,"SI_C_BD_lasso_", it ,"it.txt"))
write_tsv(res_w_HC_all, file = paste0(dir,"C_results_residuals_gene_ind_lasso_", it ,"it.txt"))
write_tsv(res_w_BD_all, file = paste0(dir, "BD_results_residuals_gene_ind_lasso_", it ,"it.txt"))

write_tsv(res_w_all, file= paste0(dir, "residuals_w_all_", it,"it.txt"))
write_tsv(mean_res_w, file= paste0(dir, "residuals_w_mean_", it,"it.txt"))
write_tsv(R2_all_long, file= paste0(dir, "R2_", it,"it.txt"))



```


```{r compare C and BD BD model}
SI_C_BD <- SI

 options(digits = 5)
SI_HC <- filter(SI_C_BD, diag=="HC") 
SI_BD <- filter(SI_C_BD,  diag=="BD")
ks.test( SI_BD$SI,  SI_HC$SI)
kstest <- ks.test( SI_BD$SI,  SI_HC$SI)


```


```{r logistic regression BD model}

SI$diag <- str_replace(SI$diag, pattern = "HC", replacement = "0")
SI$diag <- str_replace(SI$diag, pattern = "BD", replacement = "1")
SI$diag <- as.factor(SI$diag)

SI <-  cbind(Ethnicity_code = info$Ethnicity_code[match(SI$SampleID, info$SampleID)], SI)
SI <-  cbind(Sex = info$ReportedGender[match(SI$SampleID, info$SampleID)], SI)
SI <-  cbind(Age = info$Age[match(SI$SampleID, info$SampleID)], SI)
#SI <-  cbind(Study = info$Study[match(SI$SampleID, info$SampleID)], SI) #For BrainGVEX

SI_glm <- glm(diag ~ SI, data= as.data.frame(SI), family = "binomial")
SI_glm_cov <- glm(diag ~ SI + Ethnicity_code + Sex + Age,
       #         +Study, #For BrainGVEX
                  data= as.data.frame(SI), family = "binomial") #check to see which variables effect outcome

summary <- summary(SI_glm)
summary_cov <- summary(SI_glm_cov)
perf <- model_performance(SI_glm)

#show the distribution of SI values as points
SI_boxplot <- SI %>% 
  rename(Diagnosis = diag) %>% 
 ggplot( aes(y=SI, x=Diagnosis, colour= Diagnosis))+ 
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(size=0.2) +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC", "BD")) +
  labs(colour="Diagnosis") + 
  xlab("Diagnosis")+ 
  theme(legend.key.size = unit(0.5,"line"),
        legend.position = ""
        ) + 
  scale_x_discrete(breaks=c(0,1),
                    labels=c("HC", "BD")) +
  ylab("SI score")

SI_ttest <- t.test(SI ~ diag, data = SI)
SI_wilcox <- wilcox.test(SI ~ diag, data = SI)

#get pseudo R2 from the model
pseudoR2 <- nagelkerke(SI_glm)


 r2l <- result_h2l(k=0.02, p=length(BD$SampleID)/(length(HC$SampleID) + length(BD$SampleID)), r2n = pseudoR2$Pseudo.R.squared.for.model.vs.null[3] #take out the Nagelkerke
                   ) 
```

```{r save analysis BD model }
write_tsv(SI, file = paste0(dir,"SI_C_BD_wsr.txt"))
write.csv(summary$coefficients, file=paste0(dir, "glm_summary.txt"))
write.csv(summary_cov$coefficients, file=paste0(dir, "glm_covariates_summary.txt"))
write.csv(perf, file=paste0(dir, "glm_perf.txt"))
write.csv(pseudoR2$Pseudo.R.squared.for.model.vs.null, file=paste0(dir, "glm_pseudoR2.txt"))
write.csv( unlist(SI_ttest), file=paste0(dir, "SI_ttest.txt"))
write.csv( unlist(SI_wilcox), file=paste0(dir, "SI_wilcox.txt"))
write.csv(r2l, file=paste0(dir, "nagelkerke_liab.txt"))
write.csv(as.data.frame(unlist(kstest)), file = paste0(dir,"SI_kstest.txt"))
ggsave(plot= SI_boxplot, filename = paste0(dir, "SI_boxplot.pdf"), width = 43, height = 43, units = "mm")
```


Inspect the regression by ROC curves 
```{r AUC of the regression BD model}
#ROC (Receiver operating characteristic) curve
pROC_obj <- roc(SI$diag,SI$SI,
                direction= "<", #when cases have lower SI, we need to force the direction
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)
sens.ci <- ci.se(pROC_obj)


sens_ci_fix <- as_tibble(sens.ci, rownames = "spec")
sens_ci_fix$spec <- as.numeric(sens_ci_fix$spec)

pROC <- 
  ggroc(pROC_obj, size=0.5) + 
  geom_abline(slope=1, intercept = 1, linetype = "dashed", alpha=0.7, color = "black") +
  geom_ribbon(data = sens_ci_fix, 
              aes(x = spec, ymin = `2.5%`, ymax = `97.5%`), 
              fill = "#053061", alpha= 0.2) +
  xlab("Specificity") + ylab("Sensitivity")

write.csv(pROC_obj$auc, file=paste0(dir, "auc_SI_WSR.txt"))
ggsave(pROC, filename = paste0(dir, "ROC_WSR.pdf"), width = 43, height = 43, units = "mm")  
```

Calculations based on sWSR (i.e. difference from HC mean)
```{r calculate sWSR BD model}

#change directory
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda_model/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_model/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/free_coeff/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "free_coeff_lasso/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "original_gene_selection/", dataset, "/sWSR/")
#dir <- paste0("/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/", "keep_TRPT1_LMAN2L/", dataset, "/sWSR/")

#get diagnosis back in
mean_res_w_info <- mean_res_w %>% left_join(select(info, SampleID, PrimaryDiagnosis), by= "SampleID")
mean_res_w_info <- mean_res_w_info %>%  rename("wsr"= mean_res_w)

#calculate mean and sd for the HC
wsr_mean_HC <- mean_res_w_info %>% 
  filter(PrimaryDiagnosis =="HC") %>% 
  group_by(gene_id) %>% 
  mutate(mean_HC = mean(wsr),
         sd_HC = sd(wsr)) %>% 
  ungroup() %>% 
  select(gene_id, mean_HC, sd_HC) %>% 
  distinct()

#calculate sWSR-value. This is at first how many sd(HC) are the individual value from the mean HC?

sWSR_Set <- mean_res_w_info %>% 
  left_join(wsr_mean_HC, by="gene_id") %>% 
  mutate(sWSR = (wsr- mean_HC)/ sd_HC)

```


```{r SI score sWSR BD model}
#calculate SI
SI_sWSR <- sWSR_Set %>% 
 drop_na() %>%  #when all models have R2=0 for a gene, then res will be 0 and SI NA
  group_by(SampleID) %>% 
  mutate(SI = sum(abs(sWSR))/n_distinct(gene_id)) %>% #use the mean weighted scaled residual value 
  ungroup() %>% 
  select(SampleID, PrimaryDiagnosis, SI) %>%
  distinct()

SI_sWSR<- rename(SI_sWSR, "diag" = PrimaryDiagnosis)
SI_sWSR$diag <- factor(SI_sWSR$diag, levels = c("HC", "BD"))

SI_sWSR_boxplot <- 
  SI_sWSR %>% 
  rename(Diagnosis = diag) %>% 
 ggplot( aes(y=SI, x=Diagnosis, colour= Diagnosis))+ 
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(size=0.2) +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC", "BD")) +
  labs(colour="Diagnosis") + 
  xlab("Diagnosis")+ 
  theme(legend.key.size = unit(0.5,"line"),
        legend.position = ""
        ) + 
  scale_x_discrete(labels=c("HC", "BD")) +
     scale_y_continuous(
       labels = label_number(accuracy = 0.1)
       ) +
  ylab("SI score")

SI_sWSR_dens <- 
  SI_sWSR %>% 
  ggplot(aes(x=SI, colour=diag)) + 
  geom_density(size=1) +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC", "BD")) +
  #theme(legend.position = "none") +
  labs(x= "SI score", y= "Density")

ggsave(plot= SI_sWSR_boxplot, filename = paste0(dir, "SI_sWSR_boxplot.pdf"), width = 43, height = 43, units = "mm")
ggsave(plot = SI_sWSR_dens, filename = paste0(dir, "SI_density.pdf"), width = 43, height = 43, units = "mm")

```

```{r calculate ttest nagelkerke and auc BD model}

SI_sWSR_ttest <- t.test(SI ~ diag, data = SI_sWSR)
SI_sWSR_wilcox <- wilcox.test(SI ~ diag, data = SI_sWSR)

SI_sWSR$diag <- str_replace(SI_sWSR$diag, pattern = "HC", replacement = "0")
SI_sWSR$diag <- str_replace(SI_sWSR$diag, pattern = "BD", replacement = "1")
SI_sWSR$diag <- as.factor(SI_sWSR$diag)

SI_sWSR<-  cbind(Ethnicity_code = info$Ethnicity_code[match(SI_sWSR$SampleID, info$SampleID)], SI_sWSR)
SI_sWSR<-  cbind(Sex = info$ReportedGender[match(SI_sWSR$SampleID, info$SampleID)], SI_sWSR)
SI_sWSR<-  cbind(Age = info$Age[match(SI_sWSR$SampleID, info$SampleID)], SI_sWSR)
#SI_sWSR<-  cbind(Study = info$Study[match(SI_sWSR$SampleID, info$SampleID)], SI_sWSR) #For CMC-Pitt


SI_sWSR_glm <- glm(diag ~ SI, data= as.data.frame(SI_sWSR), family = "binomial")
SI_sWSR_glm_cov <- glm(diag ~ SI + Sex + Age+ Ethnicity_code,
       #         +Study  , #For BrainGVEX
                  data= as.data.frame(SI_sWSR), family = "binomial") #check to see which variables effect outcome

summary_sWSR<- summary(SI_sWSR_glm)
summary_cov_sWSR<- summary(SI_sWSR_glm_cov)
perf_sWSR<- model_performance(SI_sWSR_glm)


#get pseudo R2 from the model
pseudoR2_sWSR<- nagelkerke(SI_sWSR_glm)
r2l_sWSR <- result_h2l(k=0.02, p=length(BD$SampleID)/(length(HC$SampleID) + length(BD$SampleID)), r2n = pseudoR2_sWSR$Pseudo.R.squared.for.model.vs.null[3] #take out the Nagelkerke
                   ) 

pROC_obj <- roc(SI_sWSR$diag,SI_sWSR$SI,
                direction= "<", #when cases have lower SI, we need to force the direction
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


auc_sWSR <- pROC_obj$auc

```

```{r save sWSR results BD model}
write_tsv(sWSR_Set, paste0(dir, "sWSR.txt"))
write_tsv(SI_sWSR, file= paste0(dir, "SI_sWSR.txt"))
write.csv( unlist(SI_sWSR_ttest), file= paste0(dir,"SI_sWSR_ttest.txt"))
write.csv( unlist(SI_sWSR_wilcox), file= paste0(dir,"SI_sWSR_wilcox.txt"))
write.csv(r2l_sWSR,  file= paste0(dir,"r2l_sWSR.txt"))
write.csv(auc_sWSR[1], file=paste0(dir, "auc_sWSR.txt"))
ggsave(pROC, filename = paste0(dir, "ROC_sWSR.pdf"), width = 43, height = 43, units = "mm")  
ggsave(plot= SI_sWSR_boxplot, filename = paste0(dir, "SI_sWSR_boxplot.pdf"), width = 43, height = 43, units = "mm")
ggsave(plot = SI_sWSR_dens, filename = paste0(dir, "SI_density.pdf"), width = 43, height = 43, units = "mm")


```



```{r tstats wsr BD model}

dir <- "/Users/asbjorh/PhD/RNAseq_relative/4_cortical_sets/high_l2g_closest/lasso_cv_lambda-1se_fixed_covarites/BD_model/"


sWSR_model_BD_CMC_HBCC <- read_tsv( paste0(dir,  "CMC-HBCC/", "sWSR/sWSR.txt"))
sWSR_model_BD_GVEX_SMRI <- read_tsv( paste0(dir,  "GVEX-SMRI/", "sWSR/sWSR.txt"))
sWSR_model_BD_BipSeq_sACC<- read_tsv( paste0(dir,  "BipSeq-sACC/", "sWSR/sWSR.txt"))

sWSR_all_model_BD <- bind_rows(
  "CMC-HBCC" = sWSR_model_BD_CMC_HBCC, 
  "GVEX-SMRI" = sWSR_model_BD_GVEX_SMRI, 
  "BipSeq-sACC" = sWSR_model_BD_BipSeq_sACC, 
  .id = "Set" )

wsr_all_model_BD <- sWSR_all_model_BD %>%  select(Set, SampleID, PrimaryDiagnosis, gene_id, wsr)

wsr_all_model_BD$PrimaryDiagnosis <- factor(wsr_all_model_BD$PrimaryDiagnosis, levels = c("HC", "BD"))

#with non-parametric statistics
  stat.test_all<-
  wsr_all_model_BD %>% 
  #filter(wsr !=0) %>%  #wilcox test cannot handle many zero values well
  group_by(Set, gene_id) %>% 
  wilcox_test(data= ., formula = wsr ~ PrimaryDiagnosis, ref.group  = "HC" ) %>% 
  adjust_pvalue(method = "BH") %>% 
  mutate(y.position = 0.07) %>% 
  mutate(p.adj= round(p.adj, digits = 6))


#for Wilcoxon test, we need to calculate the direction in median difference 
#p_thr <- 0.32 #threshold for colour in tile
p_thr <- 0.15
 
wsr_median <- 
  wsr_all_model_BD %>% 
    group_by(gene_id, Set, PrimaryDiagnosis) %>% 
    mutate(median= median(wsr)) %>% 
    ungroup() %>% 
  select(Set, gene_id, PrimaryDiagnosis, median) %>%  distinct() %>%
    group_by(gene_id, Set) %>% 
  pivot_wider( names_from = PrimaryDiagnosis, values_from = median) %>% 
  mutate(median_diff = BD- HC) %>% 
  left_join(select(stat.test_all, Set, gene_id, p, p.adj), by=c("Set", "gene_id")) %>% 
  mutate(direction_diff = case_when(
    median_diff > 0 & p < p_thr ~ "pos",
    median_diff < 0 & p < p_thr ~ "neg",
    TRUE ~ "no_diff"
  )) %>%  
  ungroup() %>%  distinct()


#combine data
stat.test_all <-
  stat.test_all %>% 
  left_join(select(wsr_median, gene_id, Set, median_diff, direction_diff), by= c("Set", "gene_id")) 

stat.test_all <- stat.test_all %>% drop_na() 


stat.test_all$Set <- factor(stat.test_all$Set, levels= c("CMC-HBCC","GVEX-SMRI", "BipSeq-sACC"))

stat.test_all <- stat.test_all %>% mutate(sign = case_when(p.adj<= 0.05 ~ "FDR",
                                                           p <= 0.05 & p.adj>= 0.05 ~ "p",
                                                           TRUE ~ "none"))


stat.test_all$gene_id <- factor(stat.test_all$gene_id, levels=order.CMC_HBCC.gene_id)
stat.test_all <- stat.test_all %>% drop_na() 
statistic <- 
  ggplot(data=stat.test_all, aes(x=Set, y=gene_id)) + 
   geom_tile(aes(fill=direction_diff)) +   
         scale_fill_manual(values = c("pos" = "#F4A582", "neg" = "#92C5DE", "no_diff" = "#f7f7f7") #for wilcoxon 
                           ) +
     geom_point(data=stat.test_all, aes(size=sign, alpha=sign)) +
     scale_size_manual(values=c("FDR"=0.5, "p"=0.2, "none"= NA), guide="none") +
   scale_alpha_manual(values=c("FDR"=0.5, "p"=0.2, "none"= NA), guide="none") +
      theme(legend.title = element_blank())+
    labs(title = "" ) +
  theme_classic() +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_text(size = 5, vjust=1, angle= 45, hjust = 0.9),
          axis.text.y = element_text(size = 5, hjust=1),
          axis.ticks = element_blank(),
          legend.position = "none",
          axis.line = element_line(color='black', size = 0.2),
          panel.border = element_rect(colour = "black", fill=NA, size=0.2)
          ) + 
    #scale_y_discrete(labels= order.CMC_HBCC.gene_id) +
 # scale_x_discrete(position="top") +
    ylab("") +
  xlab("")+
     geom_hline(yintercept=seq(1.5, length(unique(stat.test_all$gene_id))+1, 1), 
             lwd=0.1, colour="gray")# Horizonal lines

  
  ggsave(statistic, filename = paste0(dir, "statistic_all_sets_model_BD.pdf"), width=4.15, height= 11.75, units="cm")
  
  
  
  
  
  plot_genes <- c("CACNA1C", "FADS2", "RASGRP1")


  
 p <- 
    wsr_all_model_BD %>%
filter( gene_id %in% plot_genes) %>%
ggplot(aes(x = PrimaryDiagnosis, y = wsr, color = PrimaryDiagnosis)) +
  theme_bw() +
  geom_boxplot(outlier.shape = NA, colour="black") +
  geom_jitter(size=0.1,width = 0.2)  +
#facet_wrap(~ factor(gene_id, levels=genes_R2$gene_id), nrow  = 1) +
  facet_grid(rows = vars(factor(gene_id, levels= plot_genes)), cols= vars(factor(Set, levels = c("CMC-HBCC", "GVEX-SMRI", "BipSeq-sACC",  "BipSeq-Amg")))) +
  scale_colour_manual(values=c(cols[2], cols[1]), labels= c("HC", "BD")) +
  scale_x_discrete(breaks=c(0,1), labels=c("HC", "BD")) +
  theme(text = element_text(size=5),
        legend.position="none",
        axis.ticks = element_blank(),
        strip.background.y  = element_rect(fill  = "#F4A582"),
         strip.background.x  = element_rect(fill  = "white")) + 
    xlab("") + ylab("") +
  coord_cartesian(ylim = c(-0.15, 0.15)) 

p_blue <- p + theme(strip.background.y = element_rect(fill  = "#92C5DE")) + coord_cartesian(ylim = c(-0.25, 0.25)) 
p_white <- p + theme(strip.background.y = element_rect(fill  = "white")) + coord_cartesian(ylim = c(-0.18, 0.15)) 

 ggsave(plot=p, path = dir, filename = "compare_res_BD_HC_redstrip_BD_model.pdf", width = 7.2, height = 10.48, units = "cm")
  ggsave(plot=p_blue, path =dir, filename = "compare_res_BD_HC_bluestrip_BD_model.pdf", width = 7.2, height = 10.48, units = "cm")
  ggsave(plot=p_white, path = dir, filename = "compare_res_BD_HC_whitestrip_BD_model.pdf", width = 7.2, height = 10.48, units = "cm")
 
```

